---
title: "Module 3: Honest DiD"
subtitle: "Sensitivity Bounds for Parallel Trends: Rambachan-Roth"
output:
  xaringan::moon_reader:
    css: [default, default-fonts]
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
---

```{css, echo=FALSE}
.remark-code, .remark-inline-code { font-size: 80%; }
.remark-slide-content { padding: 1em 2em; }
.small { font-size: 80%; }
.tiny { font-size: 65%; }
.highlight-box {
  background: #fff3e0;
  border-left: 4px solid #e65100;
  padding: 0.5em 1em;
  margin: 0.5em 0;
}
.blue-box {
  background: #e3f2fd;
  border-left: 4px solid #1565c0;
  padding: 0.5em 1em;
  margin: 0.5em 0;
}
.nav-btn {
  position: absolute;
  bottom: 12px;
  left: 40px;
  font-size: 11px;
  background: #e8eaf6;
  padding: 2px 8px;
  border-radius: 3px;
  z-index: 100;
  text-decoration: none;
  color: #1a237e;
}
.nav-btn:hover { background: #c5cae9; }
.nav-btn-br {
  position: absolute;
  bottom: 12px;
  right: 70px;
  font-size: 11px;
  background: #e8eaf6;
  padding: 2px 8px;
  border-radius: 3px;
  z-index: 100;
  text-decoration: none;
  color: #1a237e;
}
.nav-btn-br:hover { background: #c5cae9; }
.inline-btn {
  font-size: 11px;
  background: #e8eaf6;
  padding: 2px 8px;
  border-radius: 3px;
  text-decoration: none;
  color: #1a237e;
  margin-right: 6px;
  vertical-align: middle;
}
.inline-btn:hover { background: #c5cae9; }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, fig.align = "center",
  fig.width = 10, fig.height = 4.5, dpi = 150,
  message = FALSE, warning = FALSE
)
library(tidyverse)
library(fixest)
library(HonestDiD)
theme_set(theme_minimal(base_size = 14))
set.seed(42)
```

```{r make-panel, include=FALSE}
make_panel <- function(scenario = c("constant", "dynamic", "heterogeneous"),
                       n_cities = 30, n_t = 60, seed = 42) {
  scenario <- match.arg(scenario)
  set.seed(seed)
  cohorts <- tibble(
    city = 1:n_cities,
    g = c(rep(15, 8), rep(25, 8), rep(35, 8), rep(Inf, 6))
  )
  expand_grid(city = 1:n_cities, t = 1:n_t) |>
    left_join(cohorts, by = "city") |>
    mutate(
      treated = t >= g,
      cohort = case_when(g == 15 ~ "Early (g=15)", g == 25 ~ "Mid (g=25)",
                         g == 35 ~ "Late (g=35)", TRUE ~ "Never"),
      eff = case_when(
        !treated ~ 0,
        scenario == "constant"      ~ 1.0,
        scenario == "dynamic"       ~ 0.4 * (1 + 0.10 * (t - g)),
        scenario == "heterogeneous" ~ 0.4 * (1 + 0.10 * (t - g)) *
                                      (1 + 0.8 * (g == 15) - 0.8 * (g == 35))
      ),
      y = 5 + 0.05 * t + city * 0.1 + eff + rnorm(n(), 0, 0.3)
    )
}
```

```{r es-setup, include=FALSE}
# Single-cohort event study: cohort g = 25 vs the six never-treated cities,
# window rel_time in [-5, 5], clustered by city. Never-treated are pinned to
# the base period so they serve as controls at every week through the time FE.
event_fit <- function(scenario = "constant", confound = 0) {
  d <- make_panel(scenario) |>
    filter(g == 25 | !is.finite(g)) |>
    mutate(rel_time = if_else(is.finite(g), as.integer(t - 25), -1L)) |>
    filter(t >= 20, t <= 30)
  if (confound != 0)
    d <- d |> mutate(y = y + if_else(g == 25, confound * (t - 25), 0))
  feols(y ~ i(rel_time, ref = -1) | city + t, data = d, cluster = ~city)
}
tidy_es <- function(fit) {
  ev <- as.integer(sub("rel_time::", "", names(coef(fit)))); ord <- order(ev)
  list(beta = unname(coef(fit)[ord]), sigma = vcov(fit)[ord, ord], evt = ev[ord],
       numPre = sum(ev < 0), numPost = sum(ev >= 0))
}
pretest_p <- function(es) {
  pi <- which(es$evt < 0)
  W <- t(es$beta[pi]) %*% solve(es$sigma[pi, pi]) %*% es$beta[pi]
  pchisq(as.numeric(W), df = length(pi), lower.tail = FALSE)
}
bd_value <- function(sens) {                 # first M / Mbar at which CI covers 0
  col <- if ("M" %in% names(sens)) "M" else "Mbar"
  incl0 <- sens$lb <= 0 & sens$ub >= 0
  if (all(!incl0)) return(Inf)
  min(sens[[col]][incl0])
}

fit  <- event_fit("constant")
es   <- tidy_es(fit)
l1   <- basisVector(1, es$numPost)                        # first post-period
lavg <- matrix(rep(1 / es$numPost, es$numPost), ncol = 1) # average of post
ocs1 <- constructOriginalCS(es$beta, es$sigma, es$numPre, es$numPost, l_vec = l1)
p_clean <- pretest_p(es)
```

```{r hd-first, include=FALSE}
sd1 <- createSensitivityResults(es$beta, es$sigma, es$numPre, es$numPost,
         Mvec = seq(0.05, 0.6, 0.05), l_vec = l1, method = "C-LF")
rm1 <- createSensitivityResults_relativeMagnitudes(es$beta, es$sigma,
         es$numPre, es$numPost, Mbarvec = seq(0.25, 2, 0.25),
         l_vec = l1, gridPoints = 80)
```

```{r hd-avg, include=FALSE}
ocsA <- constructOriginalCS(es$beta, es$sigma, es$numPre, es$numPost, l_vec = lavg)
sdA <- createSensitivityResults(es$beta, es$sigma, es$numPre, es$numPost,
         Mvec = seq(0.05, 0.6, 0.05), l_vec = lavg, method = "C-LF")
rmA <- createSensitivityResults_relativeMagnitudes(es$beta, es$sigma,
         es$numPre, es$numPost, Mbarvec = seq(0.25, 2, 0.25),
         l_vec = lavg, gridPoints = 80)
```

```{r hd-conf, include=FALSE}
fitc <- event_fit("constant", confound = -0.10)
esc  <- tidy_es(fitc)
p_conf <- pretest_p(esc)
sd1c <- createSensitivityResults(esc$beta, esc$sigma, esc$numPre, esc$numPost,
          Mvec = seq(0.05, 0.6, 0.05), l_vec = l1, method = "C-LF")
rm1c <- createSensitivityResults_relativeMagnitudes(esc$beta, esc$sigma,
          esc$numPre, esc$numPost, Mbarvec = seq(0.25, 2, 0.25),
          l_vec = l1, gridPoints = 80)
```

```{r power-sim, include=FALSE}
# Roth's analytic pretrend-power simulation: draw betahat ~ N(beta_true, Sigma)
# with Sigma from the clean fit, a linear violation of slope delta, tau = 0.
crit <- qchisq(0.95, df = es$numPre)
Sinv <- solve(es$sigma[which(es$evt < 0), which(es$evt < 0)])
pre_idx <- which(es$evt < 0); post1 <- which(es$evt == 0)
roth_sim <- function(delta, R = 3000) {
  set.seed(100 + round(1000 * delta))
  bt <- delta * (es$evt + 1)
  dr <- MASS::mvrnorm(R, mu = bt, Sigma = es$sigma)
  wald <- apply(dr[, pre_idx, drop = FALSE], 1, function(b) t(b) %*% Sinv %*% b)
  rej <- wald > crit
  tibble(delta = delta, power = mean(rej),
         bias_uncond = mean(dr[, post1]),
         bias_cond = mean(dr[!rej, post1]))
}
power_curve <- map_dfr(seq(0, 0.15, 0.01), roth_sim)
corr_pp <- mean(cov2cor(es$sigma)[pre_idx, post1])
```

# Course Map

<table>
<tr><th>#</th><th>Module</th><th>Status</th></tr>
<tr><td>1</td><td><a href="../module-01/slides.html">TWFE Diagnosed: Goodman-Bacon and the Zoo of 2x2s</a></td><td>done</td></tr>
<tr><td>2</td><td><a href="../module-02/slides.html">Heterogeneity-Robust DiD: CS, SA, BJS, dCDH</a></td><td>done</td></tr>
<tr><td><b>3</b></td><td><b>Honest DiD: Sensitivity Bounds for Parallel Trends</b> <i>(you are here)</i></td><td>current</td></tr>
<tr><td>4</td><td><a href="../module-04/slides.html">Synthetic Control: Estimator, Inference, Variants</a></td><td>done</td></tr>
<tr><td>5</td><td><a href="../module-05/slides.html">Synthetic DiD and the Bridge from SC to DiD</a></td><td>done</td></tr>
<tr><td>6</td><td><a href="../module-06/slides.html">Causal Forest: Honest Splitting and Asymptotics</a></td><td>done</td></tr>
<tr><td>7</td><td><a href="../module-07/slides.html">Policy Learning: From HTE to Deployment Rules</a></td><td>done</td></tr>
<tr><td>8</td><td>Matrix Completion and the Modern Panel Toolbox</td><td>upcoming</td></tr>
</table>

---

# This Module

Modules 1 and 2 assumed parallel trends and fixed the *staggered-timing* bias.
This module stress-tests parallel trends itself. The tour said pre-trend tests
are weak and Rambachan-Roth bound the damage. Here is the machinery.

--

**You will be able to:**

1. Say precisely why a passing pre-trend test is weak evidence: low power and
   pre-test bias, both from Roth (2022).
2. Write the Rambachan-Roth (2023) restriction sets $\Delta^{SD}(M)$ and
   $\Delta^{RM}(\bar M)$ and say what economic story each encodes.
3. Report a **breakdown value**: how large a violation it takes to overturn a
   result.
4. Run the `HonestDiD` workflow end to end and read the sensitivity plot.

--

**Running application:** the single-cohort event study carved out of the
Module 1-2 rollout: cohort $g = 25$ versus six never-treated cities.

---

name: pretest-critique-main

# A Passing Pre-Trend Test Is Weak Evidence

.pull-left[
The event-study "pre-trend test" is a joint Wald test that the lead
coefficients are zero. Two failures, both Roth (2022):

- **Low power.** A near-linear violation is the exact shape the test cannot
  see. Below, a slope that biases the estimate is caught well under 80% of
  the time.
- **Pre-test bias.** Dropping analyses that fail the test does not clean up
  the survivors. It makes the bias **worse** (next).
]

.pull-right[
```{r power-fig, echo=FALSE, fig.height=3.6, fig.width=5}
ggplot(power_curve, aes(delta, power)) +
  geom_line(linewidth = 1, color = "#c62828") +
  geom_hline(yintercept = 0.8, linetype = "dashed") +
  geom_hline(yintercept = 0.05, linetype = "dotted") +
  labs(x = expression("linear pre-trend slope " * delta),
       y = "power of pre-trend test",
       title = "Underpowered where it matters") +
  ylim(0, 1)
```
]

.highlight-box[
Same design as our application. A slope $\delta = 0.10$ biases the naive
first-post estimate by that much yet trips the test only about
`r round(100 * power_curve$power[which.min(abs(power_curve$delta - 0.10))])`%
of the time.
]

<a href="#pretrend-power-detail" class="nav-btn">power derivation</a>
<a href="#selective-inference-detail" class="nav-btn-br">pre-test bias</a>

---

name: rr-framework-main

# The Rambachan-Roth Reframe

Stack the event-study coefficients, $\hat\beta \sim \mathcal{N}(\beta, \Sigma)$,
and split each into effect and violation:

$$\beta_{post} = \underbrace{\tau_{post}}_{\text{causal effect}} + \underbrace{\delta_{post}}_{\text{PT violation}}, \qquad \beta_{pre} = \delta_{pre}.$$

Exact PT is $\delta = 0$. **Drop it.** Assume only $\delta \in \Delta$ for a
set $\Delta$ you choose.

--

The pre-periods estimate $\delta_{pre}$ directly, so $\Delta$ links the
*observed* pre-violation to the *unobserved* post-violation. The target
$\theta = \ell' \tau_{post}$ is then **partially identified**:

$$\mathcal{S}(\beta, \Delta) = \left\{ \ell'(\beta_{post} - \delta_{post}) : \delta \in \Delta,\ \delta_{pre} = \beta_{pre} \right\}.$$

--

.blue-box[
Honesty is paid in **width**. As $\Delta$ grows the identified set widens and
the robust CI grows with it. The point estimate does not silently move; the
uncertainty does.
]

<a href="#partial-id-detail" class="nav-btn">partial identification</a>

---

name: restriction-menu-main

# The Restriction Menu

Choosing $\Delta$ *is* the method. Each choice is an economic claim about the
confounder you fear.

.small[
| $\Delta$ | Restriction | Story |
|---|---|---|
| $\Delta^{SD}(M)$ | second differences of $\delta$ bounded by $M$ | trend does not *accelerate* by more than $M$/period |
| $\Delta^{RM}(\bar M)$ | post jump $\leq \bar M \times$ worst pre jump | post violation no worse than pre, scaled |
| $\Delta^{SDPB}, \Delta^{RMB}$ | $+$ sign of $\delta_{post}$ | you know the direction of the bias |
| $\Delta^{SDM}, \Delta^{SDI}$ | $+$ monotone $\delta$ | the trend does not reverse |
]

--

.highlight-box[
Impose only what institutional knowledge defends. A sign restriction you
cannot justify buys a narrower interval you should not trust. Sign and
monotonicity *combine* with $\Delta^{SD}$ or $\Delta^{RM}$; each added
restriction shrinks the identified set.
]

---

name: delta-sd-main

# Smoothness: $\Delta^{SD}(M)$

Bound the discrete second difference of the violation:

$$\Delta^{SD}(M) = \Big\{ \delta : \big| (\delta_{s+1} - \delta_s) - (\delta_s - \delta_{s-1}) \big| \leq M \ \ \forall s \Big\}.$$

- $M = 0$: violation is exactly **linear**, extrapolated from the pre-period
  slope.
- $M > 0$: allow curvature up to $M$ per period.

--

.blue-box[
Every **linear** trend has zero second difference, so it sits in
$\Delta^{SD}(M)$ for *all* $M \geq 0$. Under $\Delta^{SD}$ the estimator
extrapolates the pre-period line into the post-period and nets it out. A purely
linear confound is something $\Delta^{SD}$ **corrects for**, not something it
flags.
]

Right when your worry is a smooth secular trend (adoption momentum, a slow
demand shift) contaminating the comparison.

<a href="#delta-sd-derivation" class="nav-btn">geometry + why linear is free</a>

---

name: delta-rm-main

# Relative Magnitudes: $\Delta^{RM}(\bar M)$

Bound the largest post jump by $\bar M$ times the largest pre jump:

$$\Delta^{RM}(\bar M) = \Big\{ \delta : |\delta_{s+1} - \delta_s| \leq \bar M \cdot \max_{s' < 0} |\delta_{s'+1} - \delta_{s'}| \ \ \forall s \geq 0 \Big\}.$$

The story: whatever moved the groups apart before treatment cannot suddenly
move them much faster after. $\bar M = 1$ means the post violation is no bigger,
period for period, than the worst wiggle already visible pre-treatment.

--

.highlight-box[
Unlike $\Delta^{SD}$, this keys on the **magnitude** of the estimated
pre-violation, so it *does* react to a linear confound: a steeper pre-trend
raises the benchmark and widens the bounds. The default when you will not assume
the trend is smooth, only that it does not change character at the treatment
date.
]

<a href="#delta-rm-derivation" class="nav-btn">the RM set, formally</a>

---

name: breakdown-main

# The Breakdown Value

Do not pick one $M$. Trace the robust CI as $M$ grows and report where the
conclusion flips:

$$M^{\ast} = \sup \Big\{ M : 0 \notin \mathcal{C}_{1-\alpha}\big(\Delta^{SD}(M)\big) \Big\},$$

the largest smoothness allowance under which the robust CI still excludes zero
(and $\bar M^{\ast}$ for $\Delta^{RM}$).

--

.blue-box[
It converts "is the effect significant?" into "**how large a PT violation would
it take to make it insignificant?**" You then judge whether a violation that
large is plausible, using the observed pre-period wiggle as the yardstick.
$\bar M^{\ast} = 1.5$ reads directly: the post violation would have to be 1.5x
the worst pre one.
]

<a href="#breakdown-derivation" class="nav-btn">reading it off the CI path</a>

---

name: robust-ci-main

# How the Robust CI Is Built

.small[
| Method | Idea | Use for |
|---|---|---|
| **FLCI** | shortest fixed-length interval, min-max over $\Delta$ | convex, centrosymmetric $\Delta$ (e.g. $\Delta^{SD}$) |
| **Conditional / C-LF** | invert an ARP moment-inequality test over $\mathcal{S}$ | non-centrosymmetric $\Delta$ (e.g. $\Delta^{RM}$) |
]

The conditional least-favorable hybrid (**C-LF**) of Andrews-Roth-Pakes is the
recommended default for $\Delta^{RM}$, where FLCI does not apply.

--

.highlight-box[
On this machine the FLCI solver is unavailable, so every run here uses
**C-LF**, which is valid for both sets. For $\Delta^{SD}$ the FLCI is usually a
touch shorter; the breakdown-value logic is identical.
]

<a href="#robust-ci-detail" class="nav-btn">FLCI vs ARP</a>

---

name: application-main

# Application: One Clean Cohort

From the Module 1-2 rollout, carve out cohort $g = 25$ (8 cities) versus the 6
never-treated cities, window $-5$ to $+5$, clustered by city.

```{r app-design, eval=FALSE}
d <- make_panel("constant") |>
  filter(g == 25 | !is.finite(g)) |>                  # one cohort + controls
  mutate(rel_time = if_else(is.finite(g), t - 25L, -1L)) |> # never-tr -> base
  filter(t >= 20, t <= 30)                            # rel_time in [-5, 5]
feols(y ~ i(rel_time, ref = -1) | city + t, data = d, cluster = ~city)
```

--

.highlight-box[
A clustered $\Sigma$ has rank at most (clusters $- 1$). With 14 clusters, a
$-10..+10$ window (20 coefficients) gives a rank-deficient $\Sigma$ and the
bounds explode. The $-5..+5$ window (10 coefficients) stays under the ceiling.
Always check the coefficient count against the cluster count.
]

<a href="#application-dgp" class="nav-btn">event-study construction</a>

---

name: event-study-main

# The Event Study

```{r es-plot, echo=FALSE, fig.height=3.8}
tibble(rel = es$evt, b = es$beta,
       se = sqrt(diag(es$sigma))) |>
  add_row(rel = -1, b = 0, se = 0) |>
  ggplot(aes(rel, b)) +
  annotate("rect", xmin = -0.5, xmax = 5.5, ymin = -Inf, ymax = Inf,
           alpha = 0.06, fill = "#1565c0") +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_vline(xintercept = -0.5, linetype = "dotted") +
  geom_pointrange(aes(ymin = b - 1.96 * se, ymax = b + 1.96 * se),
                  color = "#1565c0") +
  scale_x_continuous(breaks = -5:5) +
  labs(x = "event time (weeks since adoption)", y = "coefficient",
       title = "Flat-ish pre-period, a clear level shift after week 25")
```

Parallel trends holds by construction; the pre-test passes, with
$p = `r round(p_clean, 2)`$. But "passes" is exactly what Roth warns us to distrust.

---

name: validate-main

# First, Validate the Handoff

`HonestDiD` wants `betahat` in order (earliest pre $\dots -2$, then $0 \dots K$)
and its `sigma`. `fixest`'s `i()` already returns that order. **Confirm
`constructOriginalCS` reproduces the `feols` CI before trusting anything.**

```{r validate-code}
ocs1 <- constructOriginalCS(es$beta, es$sigma,        # numPre = 4, numPost = 6
                            es$numPre, es$numPost, l_vec = basisVector(1, es$numPost))
c(honestdid_lb = ocs1$lb, honestdid_ub = ocs1$ub,
  feols_beta   = es$beta[which(es$evt == 0)])
```

--

The `HonestDiD` "original" CI is `r sprintf("[%.3f, %.3f]", ocs1$lb, ocs1$ub)`,
matching $\hat\beta_0 \pm 1.96\,\mathrm{SE}$ for the first post-period exactly.
Ordering is right; now the sensitivity results mean something.

---

name: workflow-main

# The Workflow: Event Study In, Sensitivity Out

```{r workflow-code, eval=FALSE}
# smoothness: second differences bounded by each M
sd <- createSensitivityResults(
  betahat = es$beta, sigma = es$sigma,
  numPrePeriods = 4, numPostPeriods = 6,
  Mvec = seq(0.05, 0.6, 0.05),
  l_vec = basisVector(1, 6),          # target the FIRST post-period
  method = "C-LF")                    # FLCI solver unavailable here

# relative magnitudes: post jump <= Mbar * worst pre jump
rm <- createSensitivityResults_relativeMagnitudes(
  betahat = es$beta, sigma = es$sigma,
  numPrePeriods = 4, numPostPeriods = 6,
  Mbarvec = seq(0.25, 2, 0.25), l_vec = basisVector(1, 6))
```

Two calls, one per restriction. Each returns the robust CI at every $M$ (or
$\bar M$) in the grid. Keep the RM grid short: it is the slower of the two.

---

name: sd-result-main

# Smoothness Sensitivity, Clean Case

```{r sd-plot, echo=FALSE, fig.height=4.1}
createSensitivityPlot(sd1, ocs1) +
  labs(title = expression("Robust CI vs smoothness allowance " * M),
       subtitle = "M = 0 is the original CI; the CI widens with allowed curvature")
```

The robust CI first touches zero near $M = `r sprintf("%.2f", bd_value(sd1))`$:
that is the smoothness breakdown value $M^{\ast}$.

---

name: rm-result-main

# Relative-Magnitudes Sensitivity, Clean Case

.pull-left[
```{r rm-plot, echo=FALSE, fig.height=3.7, fig.width=5}
createSensitivityPlot_relativeMagnitudes(rm1, ocs1) +
  labs(title = expression("Robust CI vs " * bar(M)))
```
]

.pull-right[
Breakdown values, first post-period target:

.small[
| Restriction | Breakdown |
|---|---|
| $\Delta^{SD}$ | $M^{\ast} = `r sprintf("%.2f", bd_value(sd1))`$ |
| $\Delta^{RM}$ | $\bar M^{\ast} = `r sprintf("%.2f", bd_value(rm1))`$ |
]

Here the relative-magnitudes breakdown is `r sprintf("%.1f", bd_value(rm1))`, so
the post violation would have to be `r sprintf("%.1f", bd_value(rm1))`x the worst
pre-period wiggle to kill significance. Both are large relative to anything the
pre-period shows.
]

.blue-box[
**Report a number, not a verdict.** "Significant, and it survives a
parallel-trends violation `r sprintf("%.1f", bd_value(rm1))`x the pre-period one"
beats "the pre-trends looked flat."
]

---

name: target-choice-main

# The Target $\ell$ Matters

Switch $\ell$ from the first post-period to the **average** of all six. The same
data, a different estimand:

.small[
| Target $\ell$ | Original CI | $M^{\ast}$ | $\bar M^{\ast}$ |
|---|---|---|---|
| first post-period | `r sprintf("[%.2f, %.2f]", ocs1$lb, ocs1$ub)` | `r sprintf("%.2f", bd_value(sd1))` | `r sprintf("%.2f", bd_value(rm1))` |
| average of post | `r sprintf("[%.2f, %.2f]", ocsA$lb, ocsA$ub)` | `r sprintf("%.2f", bd_value(sdA))` | `r sprintf("%.2f", bd_value(rmA))` |
]

--

.highlight-box[
The average is far more fragile. Under $\Delta^{SD}$ the worst-case bias at
post-period $s$ grows like $M \cdot s^2$, so an average loading on late horizons
is exposed to compounding extrapolation. The first post-period is the most
robust target; long-run averages the least. Pick $\ell$ to match the estimand
you report, and expect later horizons to be more fragile.
]

<a href="#l-vec-detail" class="nav-btn">the compounding math</a>

---

name: confounded-main

# Confounded Case: the Restriction Must Match the Threat

Inject a differential linear trend of slope $0.10$ into the treated cohort. The
pre-test now rejects sharply, with $p = `r sprintf("%.4f", p_conf)`$. The breakdown
values split:

.small[
| Restriction | Clean | Confounded |
|---|---|---|
| $\Delta^{SD}$ | $M^{\ast} = `r sprintf("%.2f", bd_value(sd1))`$ | $M^{\ast} = `r sprintf("%.2f", bd_value(sd1c))`$ |
| $\Delta^{RM}$ | $\bar M^{\ast} = `r sprintf("%.2f", bd_value(rm1))`$ | $\bar M^{\ast} = `r sprintf("%.2f", bd_value(rm1c))`$ |
]

--

.highlight-box[
$\Delta^{RM}$ drops: the steeper pre-violation inflates the benchmark. $\Delta^{SD}$
**does not move**: a linear trend has zero second difference, so smoothness
extrapolates and nets it out. If you fear a smooth confound, $\Delta^{SD}$
*assumes it continues and corrects it* (strong). If you will not assume that,
$\Delta^{RM}$ is the honest choice.
]

<a href="#sd-vs-rm-detail" class="nav-btn">why SD is blind to a line</a>

---

name: power-bridge-main

# One Rejection Is Not a Reliable Test

In that confounded sample the pre-test rejected. So it "worked." But that is
**one draw**.

--

- Whether the test *reliably* catches a trend of slope $0.10$ is a
  repeated-sampling property.
- The power curve on slide 3 says it is caught only about
  `r round(100 * power_curve$power[which.min(abs(power_curve$delta - 0.10))])`%
  of the time.
- And conditioning on the ~`r round(100 * (1 - power_curve$power[which.min(abs(power_curve$delta - 0.10))]))`%
  that pass makes the surviving bias **worse**, because leads and lags are
  positively correlated, about $`r sprintf("%+.2f", corr_pp)`$ here.

--

.blue-box[
The gap between "this sample rejected" and "the test reliably rejects" is the
entire Roth (2022) argument, and the reason the deliverable is a breakdown value,
not a pre-test $p$. `exercise.R` reproduces both facts analytically.
]

<a href="#selective-inference-detail" class="nav-btn">conditioning is worse</a>

---

name: workflow-checklist-main

# Practitioner Checklist

.small[
| Step | What you do | Why |
|---|---|---|
| 1. Dimension | coefficient count $<$ cluster count | else $\Sigma$ is rank-deficient, bounds explode |
| 2. Validate | `constructOriginalCS` $=$ `feols` CI | catches a mis-ordered `betahat` |
| 3. Choose $\Delta$ | $\Delta^{SD}$ for smooth, $\Delta^{RM}$ for "no worse than pre" | the restriction is an economic claim |
| 4. Breakdown | report $M^{\ast}$ / $\bar M^{\ast}$, calibrate vs pre-period | one honest number |
| 5. Plot | show the sensitivity path, not a pass/fail | conveys the whole robustness curve |
]

--

At platform scale (staggered launches across marketplaces, fulfillment centers,
categories), a clean never-treated control and long pre-periods are rare, so
"how robust is this to the parallel-trends assumption?" is asked of every
DiD readout. This is the answer.

---

name: m3-interview-questions

# Interview Questions

.small[
| Question | Core of a strong answer |
|---|---|
| "The pre-trends look flat. Are we safe?" | Flat pre-trends are weak evidence: the test has low power against near-linear violations and pre-testing biases the survivors. Report a breakdown value, not a pass/fail. |
| "What is a breakdown value and how do I read it?" | The largest PT violation (in $M$ or $\bar M$) under which the robust CI still excludes zero. $\bar M^{\ast} = 1.5$: the post violation would need to be 1.5x the worst pre one. |
| "Smoothness or relative magnitudes?" | $\Delta^{SD}$ if you fear a smooth trend (it extrapolates and corrects a linear one); $\Delta^{RM}$ if you only assume the post violation is no worse than the pre. Match the restriction to the confounder. |
| "A linear confound with a clean breakdown value under $\Delta^{SD}$: trust it?" | Only if a *linear* trend is truly your worry: $\Delta^{SD}$ nets linear trends out, so its $M^{\ast}$ is blind to them. Check $\Delta^{RM}$, which reacts. |
| "How does this scale to our launches?" | Few clusters cap the event-study width (rank of $\Sigma$); validate the original CI; report breakdown values per horizon since later horizons are more fragile. |
]

---

# Going Deeper

.small[
| Paper | What it adds |
|---|---|
| Rambachan and Roth (2023), *ReStud* | The framework on these slides: the sets $\Delta$, partial identification, robust CIs, breakdown values. |
| Roth (2022), *AER: Insights* | Pre-trend tests have low power; conditioning on passing distorts inference. The exercise reproduces both. |
| Andrews, Roth and Pakes (2023), *ReStud* | The conditional moment-inequality inference (ARP / C-LF) behind the robust CIs. |
| Roth, Sant'Anna, Bilinski and Poe (2023), *J. Econometrics* | Where honest DiD sits in the modern DiD toolbox. |
| Manski (2003) | The partial-identification tradition the whole approach descends from. |
]

**Next module:** synthetic control, a different answer to "what is the
counterfactual trend?" that reweights donors instead of assuming parallelism.

**Drill:** `exercise.R` reproduces the Roth power simulation analytically and
computes breakdown values, clean vs confounded.

---

class: center, middle, inverse

# Backup Slides

---

name: pretrend-power-detail

# Backup: Roth's Power Calculation, Analytically

.small[
No panel re-simulation needed. Fit the event study **once** to get $\Sigma$, then
draw $\hat\beta \sim \mathcal{N}(\beta_{true}, \Sigma)$. Impose a linear
violation of slope $\delta$ and no true effect:

$$\beta_{true, s} = \delta \cdot (s + 1) \quad (\text{relative to base period } -1).$$

For each draw, the pre-trend Wald statistic is
$\hat\beta_{pre}' \Sigma_{pre}^{-1} \hat\beta_{pre}$; power is the share exceeding
$\chi^2_{0.95, k_{pre}}$.

```{r roth-show, eval=FALSE}
roth_sim <- function(delta, R = 3000) {
  bt <- delta * (evt + 1)                         # linear trend, tau = 0
  dr <- MASS::mvrnorm(R, mu = bt, Sigma = Sigma)
  wald <- apply(dr[, pre_idx], 1, function(b) t(b) %*% solve(Sigma_pre) %*% b)
  mean(wald > qchisq(0.95, df = numPre))          # power
}
```

Because $\hat\beta_{pre}$ is Gaussian, this is exact, not Monte Carlo error on a
DGP. The worst-case linear slope is the one the test is *least* able to see, and
that is precisely the slope that biases the post estimate most. Low power is
structural, not a small-sample accident.
]

<a href="#pretest-critique-main" class="nav-btn-br">ŌåÉ back</a>

---

name: selective-inference-detail

# Backup: Why Conditioning Makes Bias Worse

.small[
The naive first-post estimate is unbiased *for the biased estimand*
$\tau + \delta$: unconditionally $\mathbb{E}[\hat\beta_0] = \delta$ (here
$\tau = 0$). A "pre-test estimator" keeps only draws that pass, i.e. draws whose
$\hat\beta_{pre}$ landed near zero.

The violation pushes $\beta_{pre}$ *away* from zero, so a draw passes when noise
pushes it *back*. Leads and lags share the base period $-1$, so
$\mathrm{Cov}(\hat\beta_{pre}, \hat\beta_0) > 0$, about
$`r sprintf("%+.2f", corr_pp)`$ here. The same noise that shrank the leads
inflated the lag:

$$\big|\mathbb{E}[\hat\beta_0 \mid \text{pass}]\big| \; > \; \big|\mathbb{E}[\hat\beta_0]\big|.$$

| $\delta$ | naive bias | bias $\mid$ passed |
|---|---|---|
| 0.06 | `r sprintf("%.3f", power_curve$bias_uncond[which.min(abs(power_curve$delta-0.06))])` | `r sprintf("%.3f", power_curve$bias_cond[which.min(abs(power_curve$delta-0.06))])` |
| 0.10 | `r sprintf("%.3f", power_curve$bias_uncond[which.min(abs(power_curve$delta-0.10))])` | `r sprintf("%.3f", power_curve$bias_cond[which.min(abs(power_curve$delta-0.10))])` |

At $\delta = 0.10$ pre-testing roughly **doubles** the bias. Screening on
pre-trends does not rescue you; it selects the samples where the trend hid.
]

<a href="#pretest-critique-main" class="nav-btn-br">ŌåÉ back</a>

---

name: partial-id-detail

# Backup: The Identified Set, Concretely

.small[
Only the sum $\tau_{post} + \delta_{post}$ is observed. Fix the pre-period at its
estimate, $\delta_{pre} = \beta_{pre}$, and sweep $\delta_{post}$ over everything
$\Delta$ permits given that pre-period. The set of target values consistent with
the data is

$$\mathcal{S}(\beta, \Delta) = \Big\{ \ell'\big(\beta_{post} - \delta_{post}\big) : (\beta_{pre}, \delta_{post}) \in \Delta \Big\}.$$

- If $\Delta = \{0\}$ (exact PT), $\mathcal{S}$ is the point $\ell'\beta_{post}$:
  the usual event-study estimate.
- If $\Delta = \mathbb{R}^{k}$ (anything goes), $\mathcal{S} = \mathbb{R}$:
  nothing is identified without a control-trend assumption.
- The useful sets sit in between: big enough to be credible, small enough to say
  something.

Inference then covers the *set*, not a point. Two flavors: cover the whole
identified set, or cover the true parameter (the ARP conditional approach, less
conservative). `HonestDiD` does the latter.
]

<a href="#rr-framework-main" class="nav-btn-br">ŌåÉ back</a>

---

name: delta-sd-derivation

# Backup: $\Delta^{SD}$ Geometry and the Free Linear Trend

.small[
Write the violation path $\delta_{-k}, \dots, \delta_{-1}=0, \delta_0, \dots$.
The second difference $\Delta^2 \delta_s = \delta_{s+1} - 2\delta_s + \delta_{s-1}$
is the discrete curvature. $\Delta^{SD}(M)$ caps $|\Delta^2 \delta_s| \leq M$
everywhere.

Decompose any path into a **linear** part (fixed by the two base-adjacent
points) plus curvature. $\Delta^2$ annihilates the linear part, so:

- The linear component is *unrestricted* by $\Delta^{SD}(M)$ for any $M$. The
  identified set is centered on the linear extrapolation of the pre-trend.
- $M$ controls only how far the post path may bow away from that line. The
  worst-case deviation at post-period $s$ grows like $M \cdot s^2$ (curvature
  integrated twice), which is why late horizons blow up first.

Consequence: shifting the whole sample by a linear confound moves the estimate
and the extrapolation **together**, leaving $M^{\ast}$ unchanged (slide
[confounded case](#confounded-main)). $\Delta^{SD}$ answers "how much curvature
can I tolerate?", never "is there a line?".
]

<a href="#delta-sd-main" class="nav-btn-br">ŌåÉ back</a>

---

name: delta-rm-derivation

# Backup: $\Delta^{RM}$, Formally

.small[
Let the first difference be $\Delta \delta_s = \delta_{s+1} - \delta_s$ (the
period-to-period change in the violation). Define the worst pre-period change

$$D_{pre} = \max_{s' < 0} |\Delta \delta_{s'}|,$$

which the pre-period coefficients estimate. $\Delta^{RM}(\bar M)$ requires every
post-period change to obey

$$|\Delta \delta_s| \leq \bar M \cdot D_{pre}, \qquad s \geq 0.$$

- The set is a scaled cone anchored to the *observed* pre-period wiggle, not an
  absolute magnitude, hence "relative."
- It is **not centrosymmetric**, so FLCI does not apply and inference uses the
  ARP conditional / C-LF construction.
- A steeper pre-trend raises $D_{pre}$, enlarges the set, and lowers
  $\bar M^{\ast}$: the restriction reacts to exactly the linear confound that
  $\Delta^{SD}$ ignores.

Reach for $\Delta^{RM}$ when you are unwilling to assume smoothness but will
assume the post-period trend is no wilder than the pre-period one.
]

<a href="#delta-rm-main" class="nav-btn-br">ŌåÉ back</a>

---

name: breakdown-derivation

# Backup: Reading the Breakdown Value Off the CI Path

.small[
For each $M$, `HonestDiD` returns a robust CI $[\text{lb}(M), \text{ub}(M)]$.
Both endpoints move monotonically outward in $M$ (a bigger $\Delta$ can only
enlarge the identified set). Define the breakdown value as the first $M$ at
which the CI covers zero:

$$M^{\ast} = \min\{ M : \text{lb}(M) \leq 0 \leq \text{ub}(M) \}.$$

Practically, run a grid and find where $\text{lb}(M)$ crosses zero. In the clean
case, $\text{lb}$ falls from `r sprintf("%.2f", sd1$lb[1])` at $M = 0.05$ through
zero near $M = `r sprintf("%.2f", bd_value(sd1))`$:

```{r bd-show, echo=FALSE}
sd1 |> transmute(M, lb = round(lb, 3), ub = round(ub, 3),
                 excludes_0 = !(lb <= 0 & ub >= 0)) |> head(10)
```

Report $M^{\ast}$ with its yardstick: "the effect survives curvature up to
$M^{\ast}$, versus a pre-period second difference of at most (observed value)."
A breakdown value without a scale to compare it to is not yet an argument.
]

<a href="#breakdown-main" class="nav-btn-br">ŌåÉ back</a>

---

name: robust-ci-detail

# Backup: FLCI vs the ARP Conditional Approach

.small[
**FLCI** (fixed-length CI). Restrict to affine estimators $a + v'\hat\beta$ and
choose $v$ to minimize the CI length subject to worst-case coverage over
$\delta \in \Delta$. This is a tractable convex program **only when $\Delta$ is
convex and centrosymmetric**, i.e. $\delta \in \Delta \Rightarrow -\delta \in \Delta$,
which $\Delta^{SD}(M)$ satisfies but $\Delta^{RM}(\bar M)$ does not. FLCI gives the
shortest honest interval when it applies.

**ARP conditional (Andrews-Roth-Pakes).** Cast $\delta \in \Delta$ as linear
moment inequalities and invert a test of $H_0: \theta = \theta_0$ over a grid of
$\theta_0$. The conditional test corrects for the fact that which inequalities
bind is data-dependent. The **least-favorable hybrid (C-LF)** blends a
least-favorable first stage with the conditional test for power. It is valid for
*any* polyhedral $\Delta$, so it covers $\Delta^{RM}$ and the combined sets.

Rule of thumb: $\Delta^{SD}$ alone -> FLCI (shortest); anything with relative
magnitudes, signs, or monotonicity -> C-LF. Here FLCI is unavailable, so C-LF is
used throughout, at a small length cost for $\Delta^{SD}$.
]

<a href="#robust-ci-main" class="nav-btn-br">ŌåÉ back</a>

---

name: application-dgp

# Backup: Building the Event Study

.small[
```{r es-dgp-show, eval=FALSE}
event_fit <- function(scenario = "constant", confound = 0) {
  d <- make_panel(scenario) |>
    filter(g == 25 | !is.finite(g)) |>                     # one cohort + never-tr
    mutate(rel_time = if_else(is.finite(g), as.integer(t - 25), -1L)) |>
    filter(t >= 20, t <= 30)                               # rel_time in [-5, 5]
  if (confound != 0)                                       # optional PT violation
    d <- d |> mutate(y = y + if_else(g == 25, confound * (t - 25), 0))
  feols(y ~ i(rel_time, ref = -1) | city + t, data = d, cluster = ~city)
}
tidy_es <- function(fit) {                                 # HonestDiD-ready pieces
  ev <- as.integer(sub("rel_time::", "", names(coef(fit)))); ord <- order(ev)
  list(beta = unname(coef(fit)[ord]), sigma = vcov(fit)[ord, ord],
       evt = ev[ord], numPre = sum(ev < 0), numPost = sum(ev >= 0))
}
```

- Never-treated cities get `rel_time = -1` so they land in the base category and
  serve as controls at every week through the time fixed effects.
- `filter(t >= 20, t <= 30)` matches the calendar window to $[-5, +5]$ event
  time for the $g = 25$ cohort, so the time FE are estimated only where the
  treated cohort has an event-time coefficient.
- `order(ev)` guarantees the (pre earliest $\dots$ post) ordering `HonestDiD`
  expects. This is the step to validate against the `feols` CI.
]

<a href="#application-main" class="nav-btn-br">ŌåÉ back</a>

---

name: l-vec-detail

# Backup: Why the Average Target Is Fragile

.small[
The target is $\theta = \ell' \tau_{post}$. Two natural choices:

- **First post-period:** $\ell = (1, 0, 0, 0, 0, 0)'$. Under $\Delta^{SD}(M)$
  the worst-case bias is bounded by $M \cdot 1^2$: only one period of
  extrapolation.
- **Average of post:** $\ell = (\tfrac16, \dots, \tfrac16)'$. The bias at
  horizon $s$ is bounded by $\sim M \cdot s^2$, so the average is
  $\frac{M}{6}\sum_{s=0}^{5} s^2 \approx M \cdot 9$: nine times the exposure.

That is why the same panel gives $M^{\ast} = `r sprintf("%.2f", bd_value(sd1))`$
for the first period but $M^{\ast} = `r sprintf("%.2f", bd_value(sdA))`$ for the
average, and $\bar M^{\ast}$ falls from `r sprintf("%.2f", bd_value(rm1))` to
`r sprintf("%.2f", bd_value(rmA))`. The average is not "wrong": it answers a
different, later-horizon question that is intrinsically harder to defend against
trend extrapolation.

Practical rule: report the breakdown value **for the horizon you will act on**.
If the decision hinges on the long-run average, its fragility is real
information, not a nuisance.
]

<a href="#target-choice-main" class="nav-btn-br">ŌåÉ back</a>

---

name: sd-vs-rm-detail

# Backup: The Same Confound, Two Verdicts

.small[
The injected confound is $0.10 \cdot (t - 25)$ added to the treated cohort: a
pure line through the treatment date. In event-study coordinates it shifts every
coefficient by a linear amount.

| | clean | confounded |
|---|---|---|
| pre-test $p$ | `r sprintf("%.2f", p_clean)` | `r sprintf("%.4f", p_conf)` |
| $M^{\ast}$ under $\Delta^{SD}$ | `r sprintf("%.2f", bd_value(sd1))` | `r sprintf("%.2f", bd_value(sd1c))` |
| $\bar M^{\ast}$ under $\Delta^{RM}$ | `r sprintf("%.2f", bd_value(rm1))` | `r sprintf("%.2f", bd_value(rm1c))` |

- **Smoothness does not react.** A line has zero second difference, so it is
  inside $\Delta^{SD}(0)$. The estimator extrapolates the pre-period line into
  the post and subtracts it; the estimate and the correction move together and
  $M^{\ast}$ is untouched. $\Delta^{SD}$ is *defined* to be blind to a linear
  confound.
- **Relative magnitudes does react.** The confound enlarges the worst pre-period
  jump $D_{pre}$, so $\Delta^{RM}(\bar M)$ permits a proportionally larger post
  violation and the bounds widen: $\bar M^{\ast}$ falls
  `r sprintf("%.2f", bd_value(rm1))` $\to$ `r sprintf("%.2f", bd_value(rm1c))`.

Neither is "more correct." They encode different beliefs. The takeaway: never
read a breakdown value without stating which $\Delta$ produced it and whether
that $\Delta$ can even see the confound you fear.
]

<a href="#confounded-main" class="nav-btn-br">ŌåÉ back</a>
