---
title: "Module 1: TWFE Diagnosed"
subtitle: "Goodman-Bacon and the Zoo of 2x2s"
output:
  xaringan::moon_reader:
    css: [default, default-fonts]
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
---

```{css, echo=FALSE}
.remark-code, .remark-inline-code { font-size: 80%; }
.remark-slide-content { padding: 1em 2em; }
.small { font-size: 80%; }
.tiny { font-size: 65%; }
.highlight-box {
  background: #fff3e0;
  border-left: 4px solid #e65100;
  padding: 0.5em 1em;
  margin: 0.5em 0;
}
.blue-box {
  background: #e3f2fd;
  border-left: 4px solid #1565c0;
  padding: 0.5em 1em;
  margin: 0.5em 0;
}
.nav-btn {
  position: absolute;
  bottom: 12px;
  left: 40px;
  font-size: 11px;
  background: #e8eaf6;
  padding: 2px 8px;
  border-radius: 3px;
  z-index: 100;
  text-decoration: none;
  color: #1a237e;
}
.nav-btn:hover { background: #c5cae9; }
.nav-btn-br {
  position: absolute;
  bottom: 12px;
  right: 70px;
  font-size: 11px;
  background: #e8eaf6;
  padding: 2px 8px;
  border-radius: 3px;
  z-index: 100;
  text-decoration: none;
  color: #1a237e;
}
.nav-btn-br:hover { background: #c5cae9; }
.inline-btn {
  font-size: 11px;
  background: #e8eaf6;
  padding: 2px 8px;
  border-radius: 3px;
  text-decoration: none;
  color: #1a237e;
  margin-right: 6px;
  vertical-align: middle;
}
.inline-btn:hover { background: #c5cae9; }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, fig.align = "center",
  fig.width = 10, fig.height = 4.5, dpi = 150,
  message = FALSE, warning = FALSE
)
library(tidyverse)
library(fixest)
library(bacondecomp)
theme_set(theme_minimal(base_size = 14))
set.seed(42)
```

```{r make-panel, include=FALSE}
make_panel <- function(scenario = c("constant", "dynamic", "heterogeneous"),
                       n_cities = 30, n_t = 60, seed = 42) {
  scenario <- match.arg(scenario)
  set.seed(seed)
  cohorts <- tibble(
    city = 1:n_cities,
    g = c(rep(15, 8), rep(25, 8), rep(35, 8), rep(Inf, 6))
  )
  expand_grid(city = 1:n_cities, t = 1:n_t) |>
    left_join(cohorts, by = "city") |>
    mutate(
      treated = t >= g,
      cohort = case_when(g == 15 ~ "Early (g=15)", g == 25 ~ "Mid (g=25)",
                         g == 35 ~ "Late (g=35)", TRUE ~ "Never"),
      eff = case_when(
        !treated ~ 0,
        scenario == "constant"      ~ 1.0,
        scenario == "dynamic"       ~ 0.4 * (1 + 0.10 * (t - g)),
        scenario == "heterogeneous" ~ 0.4 * (1 + 0.10 * (t - g)) *
                                      (1 + 0.8 * (g == 15) - 0.8 * (g == 35))
      ),
      y = 5 + 0.05 * t + city * 0.1 + eff + rnorm(n(), 0, 0.3)
    )
}
panel <- make_panel("dynamic")
```

# Course Map

<table>
<tr><th>#</th><th>Module</th><th>Status</th></tr>
<tr><td><b>1</b></td><td><b>TWFE Diagnosed: Goodman-Bacon and the Zoo of 2x2s</b> <i>(you are here)</i></td><td>current</td></tr>
<tr><td>2</td><td><a href="../module-02/slides.html">Heterogeneity-Robust DiD: CS, SA, BJS, dCDH</a></td><td>done</td></tr>
<tr><td>3</td><td><a href="../module-03/slides.html">Honest DiD: Sensitivity Bounds for Parallel Trends</a></td><td>done</td></tr>
<tr><td>4</td><td><a href="../module-04/slides.html">Synthetic Control: Estimator, Inference, Variants</a></td><td>done</td></tr>
<tr><td>5</td><td><a href="../module-05/slides.html">Synthetic DiD and the Bridge from SC to DiD</a></td><td>done</td></tr>
<tr><td>6</td><td><a href="../module-06/slides.html">Causal Forest: Honest Splitting and Asymptotics</a></td><td>done</td></tr>
<tr><td>7</td><td><a href="../module-07/slides.html">Policy Learning: From HTE to Deployment Rules</a></td><td>done</td></tr>
<tr><td>8</td><td><a href="../module-08/slides.html">Matrix Completion and the Modern Panel Toolbox</a></td><td>done</td></tr>
</table>

---

# This Module

The experimentation-refresher tour showed *that* TWFE breaks under staggered
adoption. This module shows *exactly how*, with the full Goodman-Bacon (2021)
decomposition.

--

**You will be able to:**

1. State the TWFE estimand under staggered timing as a weighted average of
   all possible 2x2 DiD comparisons.
2. Derive and compute the Bacon weights by hand, and replicate
   `bacondecomp::bacon()` to machine precision.
3. Explain the forbidden comparison and show when it flips signs.
4. Say precisely when TWFE is still fine, and defend that in an interview.

--

**Running application:** staggered rollout of a driver zone-notification
feature across 30 cities, 60 weeks, three adoption cohorts plus 6
never-treated cities.

---

name: twfe-refresher-main

# The Estimator Under the Microscope

The two-way fixed effects (TWFE) regression:

$$y_{it} = \alpha_i + \lambda_t + \beta^{DD} D_{it} + \varepsilon_{it}$$

- $\alpha_i$: unit fixed effects. $\lambda_t$: period fixed effects.
- $D_{it} \in \{0,1\}$: unit $i$ treated at time $t$.

--

.pull-left[
**With 2 groups, 2 periods** this *is* the DiD estimator:

$$\hat\beta^{DD} = \Delta \bar y_{\text{treat}} - \Delta \bar y_{\text{ctrl}}$$

Unbiased for the ATT under parallel trends. No controversy.
]

.pull-right[
**With staggered timing** it is still a single number, but *which* number?

By Frisch-Waugh-Lovell:

$$\hat\beta^{DD} = \frac{\widehat{Cov}(y_{it}, \tilde D_{it})}{\widehat{Var}(\tilde D_{it})}$$

where $\tilde D_{it}$ is $D_{it}$ residualized on both fixed effects.
]

--

.small[
.highlight-box[
Everything wrong with TWFE under staggered adoption hides inside
$\tilde D_{it}$: units treated *the whole panel* or *never* have no
variation left, and mid-panel switchers dominate the variance.
]
]

<a href="#fwl-detail" class="nav-btn">FWL detail</a>

---

name: staggered-dgp-main

# The Application: Staggered Zone-Notification Rollout

30 cities, 60 weeks. Cohorts adopt at week 15, 25, 35; six cities never do.
Outcome $y$: log completed trips (read: engagement or revenue per customer
in a retail setting).

```{r dgp-plot, echo=FALSE, fig.height=2.7}
panel |>
  group_by(cohort, t) |>
  summarise(y = mean(y), .groups = "drop") |>
  ggplot(aes(t, y, color = cohort)) +
  geom_line(linewidth = 1) +
  geom_vline(xintercept = c(15, 25, 35), linetype = "dashed", alpha = 0.4) +
  labs(x = "Week", y = "Mean log trips", color = NULL,
       title = "Dynamic scenario: effects grow with time since adoption") +
  theme(legend.position = "bottom")
```

.tiny[
Units here are cities, which keeps cross-unit interference plausibly small.
If units were sellers or customers inside one marketplace (shared
inventory, cannibalization, reallocation), spillovers would contaminate
not-yet-treated controls, and no staggered-DiD estimator fixes that; unit
choice is the first identification decision.
]

<a href="#staggered-dgp" class="nav-btn">DGP code</a>

---

name: scenarios-main

# Three Heterogeneity Scenarios, One DGP

Same panel structure; only the treatment-effect function changes.

.small[
| Scenario | Effect for cohort $g$ at time $t$ | Interpretation |
|---|---|---|
| `constant` | $1.0$ | one-time level shift, same for everyone |
| `dynamic` | $0.4\,\lbrack 1 + 0.10 (t-g)\rbrack$ | effect grows with exposure, same across cohorts |
| `heterogeneous` | $0.4\,\lbrack 1 + 0.10 (t-g)\rbrack\,\lbrack 1 + 0.8\cdot\mathbb{1}(g{=}15) - 0.8\cdot\mathbb{1}(g{=}35)\rbrack$ | early adopters gain most |
]

--

```{r scenario-table, echo=FALSE}
res <- map_dfr(c("constant", "dynamic", "heterogeneous"), function(s) {
  p <- make_panel(s)
  tibble(
    Scenario = s,
    `True ATT` = mean(p$eff[p$treated]),
    `TWFE estimate` = coef(feols(y ~ treated | city + t, data = p))[[1]]
  )
})
res |>
  mutate(`Bias (%)` = 100 * (`TWFE estimate` / `True ATT` - 1)) |>
  mutate(across(where(is.numeric), ~ round(.x, 3))) |>
  knitr::kable(format = "html")
```

--

.highlight-box[
TWFE recovers the constant effect, loses 40% of the dynamic one, and 70%
once cohorts differ. Nothing about the code changed. The estimator did this.
]

---

name: bacon-thm-main

# Goodman-Bacon (2021): The Theorem

With $K$ timing groups (adoption dates), the TWFE coefficient is **exactly**

$$\hat\beta^{DD} = \sum_{k \neq U} s_{kU}\, \hat\beta_{kU} + \sum_{k} \sum_{l > k} \left[ s_{kl}^{k}\, \hat\beta_{kl}^{k} + s_{kl}^{l}\, \hat\beta_{kl}^{l} \right]$$

a weighted average of **every possible 2x2 DiD** in the data, with weights
$s \geq 0$ summing to 1.

--

The three species in the zoo:

.small[
| 2x2 | Treatment group | Control group | Window |
|---|---|---|---|
| $\hat\beta_{kU}$ | cohort $k$ | never-treated $U$ | full panel |
| $\hat\beta_{kl}^{k}$ | earlier cohort $k$ | later cohort $l$, *not yet treated* | $t < g_l$ |
| $\hat\beta_{kl}^{l}$ | later cohort $l$ | earlier cohort $k$, **already treated** | $t \geq g_k$ |
]

--

.blue-box[
The first two are legitimate DiDs. The third uses already-treated units as
controls: the **forbidden comparison**.
]

<a href="#bacon-derivation" class="nav-btn">derivation</a>

---

name: bacon-timing-main

# Who Gets Compared to Whom

```{r timing-diagram, echo=FALSE, fig.height=4.8}
windows <- tribble(
  ~comparison, ~role, ~t0, ~t1, ~grp,
  "Early (15) vs Never",       "treatment", 1, 60, "Early",
  "Early (15) vs Never",       "control",   1, 60, "Never",
  "Early (15) vs Late (35),\nlate is control", "treatment", 1, 34, "Early",
  "Early (15) vs Late (35),\nlate is control", "control",   1, 34, "Late",
  "Late (35) vs Early (15),\nFORBIDDEN",       "treatment", 15, 60, "Late",
  "Late (35) vs Early (15),\nFORBIDDEN",       "control",   15, 60, "Early"
) |>
  mutate(ypos = as.numeric(factor(paste(comparison, role))))
adopt <- tibble(grp = c("Early", "Late"), g = c(15, 35))
windows |>
  left_join(adopt, by = "grp") |>
  ggplot() +
  geom_segment(aes(x = t0, xend = t1, y = ypos, yend = ypos, color = role),
               linewidth = 5, alpha = 0.7) +
  geom_point(aes(x = g, y = ypos), shape = 4, size = 4, stroke = 2,
             na.rm = TRUE) +
  geom_text(aes(x = t0, y = ypos + 0.28, label = grp), hjust = 0, size = 3.4) +
  facet_wrap(~comparison, ncol = 1, scales = "free_y") +
  scale_y_continuous(expand = expansion(add = 0.7)) +
  scale_color_manual(values = c(treatment = "#1565c0", control = "#e65100")) +
  labs(x = "Week", y = NULL, color = NULL,
       title = "X marks adoption. The forbidden 2x2's control adopts inside its own window.") +
  theme(axis.text.y = element_blank(), legend.position = "bottom")
```

---

name: bacon-weights-main

# The Weights

Let $n_k$ be group sizes, $n_{kl} = n_k / (n_k + n_l)$ relative sizes, and
$\bar D_k$ the share of the panel cohort $k$ spends treated. Then:

.small[
$$s_{kU} \propto (n_k + n_U)^2 \; n_{kU}(1 - n_{kU}) \; \bar D_k (1 - \bar D_k)$$

$$s_{kl}^{k} \propto \left[(n_k + n_l)(1 - \bar D_l)\right]^2 \; n_{kl}(1 - n_{kl}) \; \frac{\bar D_k - \bar D_l}{1 - \bar D_l} \cdot \frac{1 - \bar D_k}{1 - \bar D_l}$$

$$s_{kl}^{l} \propto \left[(n_k + n_l)\bar D_k\right]^2 \; n_{kl}(1 - n_{kl}) \; \frac{\bar D_l}{\bar D_k} \cdot \frac{\bar D_k - \bar D_l}{\bar D_k}$$
]

--

**What drives a weight up:**

- Bigger subsamples: the $(n_k + n_l)^2$ and $n_{kl}(1-n_{kl})$ terms.
- Treatment-share variance $\bar D(1-\bar D)$: cohorts treated **near the
  middle of the panel** get the most weight. Units treated very early or
  very late contribute little identifying variance.

--

.highlight-box[
Weights come from the *design* (sizes and timing), not from effect
precision. TWFE decides what to estimate based on when your PM shipped.
]

<a href="#weights-detail" class="nav-btn">where these come from</a>

---

name: forbidden-main

# The Forbidden Comparison, Up Close

Late cohort (g=35) as treatment, Early cohort (g=15) as control, window
$t \geq 15$. Both cohorts have strictly positive, growing effects.

```{r forbidden-plot, echo=FALSE, fig.height=4.2}
panel |>
  filter(cohort %in% c("Early (g=15)", "Late (g=35)"), t >= 15) |>
  group_by(cohort, t) |>
  summarise(y = mean(y), .groups = "drop") |>
  ggplot(aes(t, y, color = cohort)) +
  annotate("rect", xmin = 35, xmax = 60, ymin = -Inf, ymax = Inf,
           alpha = 0.08, fill = "#e65100") +
  geom_line(linewidth = 1) +
  geom_vline(xintercept = 35, linetype = "dashed") +
  annotate("text", x = 47, y = 8.2, label = "late cohort's post window", size = 3.5) +
  scale_color_manual(values = c("Early (g=15)" = "#e65100", "Late (g=35)" = "#1565c0")) +
  labs(x = "Week", y = "Mean log trips", color = NULL,
       title = "The 'control' is on its own treatment-effect trend") +
  theme(legend.position = "bottom")
```

--

The early cohort's *growing* effect is part of the control trend and gets
**subtracted**. This 2x2 estimates $\text{ATT}_{late} - \Delta\text{ATT}_{early}$,
not an ATT.

<a href="#forbidden-detail" class="nav-btn">the algebra</a>

---

name: bacon-practice-main

# `bacondecomp::bacon()` in Practice

```{r bacon-run, results='hide'}
library(bacondecomp)
df <- panel |> # bacon() wants a plain data.frame, numeric treatment
  transmute(city, t, y, treat = as.numeric(treated)) |>
  as.data.frame()
bd <- bacon(y ~ treat, data = df, id_var = "city", time_var = "t")
```

--

```{r bacon-table}
bd |> group_by(type) |> # aggregate the 9 2x2s to the 3 species
  summarise(avg_estimate = weighted.mean(estimate, weight),
            weight = sum(weight)) |>
  mutate(across(where(is.numeric), ~ round(.x, 3)))
```

--

.highlight-box[
A third of the total weight sits on forbidden comparisons averaging 0.08,
against legitimate comparisons averaging around 1.1. That is the whole
attenuation story in one table.
]

---

name: bacon-scatter-main

# The Diagnostic Plot: Weight vs Estimate

```{r bacon-scatter, echo=FALSE, fig.height=4.6}
truth <- mean(panel$eff[panel$treated])
twfe_est <- coef(feols(y ~ treated | city + t, data = panel))[[1]]
bd |>
  ggplot(aes(weight, estimate, color = type, shape = type)) +
  geom_hline(yintercept = twfe_est, linetype = "dashed") +
  geom_hline(yintercept = truth, linetype = "dotted") +
  geom_point(size = 4) +
  annotate("text", x = 0.155, y = twfe_est - 0.07, label = "TWFE", size = 3.5) +
  annotate("text", x = 0.155, y = truth + 0.07, label = "true ATT", size = 3.5) +
  scale_color_manual(values = c("#1565c0", "#e65100", "#2e7d32")) +
  labs(x = "Bacon weight", y = "2x2 DiD estimate", color = NULL, shape = NULL,
       title = "Each point is one 2x2; TWFE is their weighted average") +
  theme(legend.position = "bottom")
```

Every applied seminar wants this plot before believing a staggered DiD.

---

name: bacon-validate-main

# The Decomposition Is Exact

```{r validate}
c(weighted_sum = sum(bd$estimate * bd$weight),
  twfe         = coef(feols(y ~ treated | city + t, data = panel))[[1]])
```

--

Identical to machine precision. This is an algebraic identity about OLS,
**not an approximation and not an estimator of anything new**.

--

.blue-box[
**Reading it as an estimand decomposition:** under parallel trends,

$$\text{plim}\;\hat\beta^{DD} = \text{VWATT} - \Delta\text{ATT}$$

a *variance-weighted* ATT minus a term collecting **within-cohort effect
changes** that forbidden comparisons subtract. Constant effects kill
$\Delta\text{ATT}$; dynamics feed it.
]

<a href="#estimand-detail" class="nav-btn">full estimand decomposition</a>

---

name: twfe-ok-main

# When TWFE Is Perfectly Fine

Run the identical decomposition on the `constant` scenario:

```{r constant-bacon, echo=FALSE}
p_const <- make_panel("constant")
df_c <- p_const |>
  transmute(city, t, y, treat = as.numeric(treated)) |>
  as.data.frame()
junk <- capture.output( # suppress bacon()'s auto-printed summary
  bd_c <- bacon(y ~ treat, data = df_c, id_var = "city", time_var = "t"))
bd_c |> group_by(type) |>
  summarise(avg_estimate = weighted.mean(estimate, weight),
            weight = sum(weight)) |>
  mutate(across(where(is.numeric), ~ round(.x, 3))) |>
  knitr::kable(format = "html")
```

--

Weights are unchanged (they depend only on the design). But now **every 2x2
estimates the same 1.0**, so the weighting is harmless.

--

**TWFE is unbiased for the ATT when:**

1. Effects are homogeneous across cohorts *and* constant in event time; or
2. Adoption is *not* staggered (single treatment date, 2x2 world); or
3. You are willing to interpret VWATT as the estimand and effects are
   stable over time.

.highlight-box[
The interview answer is not "TWFE is broken". It is: "TWFE estimates a
variance-weighted average of 2x2s; here is when that coincides with the
ATT, and here is the diagnostic I would run."
]

---

name: negative-weights-main

# From Attenuation to Sign Reversal

Bacon weights are non-negative *on the 2x2 estimates*. The scary headlines
("negative weights") are about the implied weights **on the underlying
cohort-time effects** $ATT(g,t)$ inside those 2x2s.

--

- Each forbidden $\hat\beta_{kl}^{l}$ contains $-\Delta ATT_k$: early-cohort
  effect *growth* enters with a negative sign.
- Our panel resists a full flip because 6 never-treated cities anchor half
  the weight on clean comparisons. **Remove them** (everyone eventually
  adopts, common at a company that ships everywhere) and forbidden 2x2s take
  over.

--

```{r sign-demo, echo=TRUE}
set.seed(42) # no never-treated group: 3 cohorts of 10, everyone adopts
flip <- expand_grid(city = 1:30, t = 1:60) |>
  left_join(tibble(city = 1:30, g = rep(c(15, 25, 35), each = 10)), by = "city") |>
  mutate(treated = t >= g,
         eff = if_else(treated, 0.2 + 0.15 * (t - g), 0), # mild dynamics
         y = 5 + 0.05 * t + city * 0.1 + eff + rnorm(n(), 0, 0.3))
c(true_ATT = mean(flip$eff[flip$treated]),
  TWFE     = coef(feols(y ~ treated | city + t, data = flip))[[1]])
```

--

Every $ATT(g,t) > 0$, TWFE **negative**. de Chaisemartin and D'Haultfoeuille
(2020) formalize the implicit weights behind this; that is Module 2
territory.

---

name: design-rollout-main

# Design the Rollout: Reserve a Never-Treated Slice

The sign-reversal demo above is not just a diagnosis, it is an argument you
can make **before** a staggered rollout ships. If you are consulted on the
design, argue to hold out a slice of markets or cohorts, the panel analogue
of a global holdout in an A/B test.

--

- A platform-wide launch cannot be a customer-level experiment, but **which
  cities wait** is still a design choice you can make.
- Even a small never-treated set restores clean comparisons for every
  estimator in this course, not just TWFE: it anchors weight away from
  forbidden 2x2s here, and it is the control group Callaway-Sant'Anna and
  BJS need in Module 2.
- The quantified case is the demo above: with 6 never-treated cities the
  panel resists a full sign flip; remove them and it does not.

--

.blue-box[
"Ship everywhere" is an ops default, not a technical requirement. A held-out
slice costs some short-run coverage and buys identification for the entire
toolbox that follows.
]

---

name: diagnostics-main

# Practitioner Checklist for Any Staggered Rollout

Before reporting a staggered TWFE estimate:

.small[
| Step | Tool | What you learn |
|---|---|---|
| 1. Map the cohorts | `count(g)` | how much never-treated glue you have |
| 2. Bacon decomposition | `bacondecomp::bacon()` | weight on forbidden comparisons |
| 3. Weight-vs-estimate scatter | previous slide | which 2x2s drive the coefficient |
| 4. Compare vs robust estimator | CS / SA / BJS (Module 2) | how much the answer moves |
| 5. Event-study with robust estimator | Module 2 | dynamics, pre-trends |
]

--

.blue-box[
**Rules of thumb.** Forbidden weight near zero and effects plausibly stable:
TWFE is probably fine, say why. Forbidden weight 30%+ or visible dynamics:
robust estimators are not optional.
]

--

At platform scale (staggered feature launches across marketplaces, fulfillment
centers, product categories), staggering is the *default*, not the edge case.

---

name: m1-interview-questions

# Interview Questions

.small[
| Question | Core of a strong answer |
|---|---|
| "We rolled a feature out city by city and ran TWFE. What could go wrong?" | Staggered timing makes TWFE a weighted average of 2x2s including already-treated controls; with dynamic effects it is biased toward zero or beyond. Run the Bacon decomposition. |
| "What exactly is a forbidden comparison?" | A 2x2 using an earlier-treated cohort as control during its own post period; its effect *changes* get subtracted from the later cohort's estimate. |
| "When would you still trust TWFE?" | Non-staggered designs; effects constant in event time and across cohorts; or when forbidden-comparison weight is negligible and you can defend effect stability. |
| "Can TWFE flip the sign of a true positive effect?" | Yes: strong effect dynamics make forbidden 2x2s increasingly negative, and they can dominate. Demonstrated with all $ATT(g,t) > 0$ and TWFE $< 0$. |
| "What determines the Bacon weights?" | Group sizes and treatment-share variance $\bar D(1-\bar D)$: mid-panel cohorts dominate. Design, not effect size or precision. |
| "Ops staggered the launch so every city eventually got it. That is even better than a holdout, right?" | No: with no never-treated slice, late cohorts get compared against already-treated cities, and TWFE can even flip the sign of a true positive effect (this deck's demo). Reserve a never-treated holdout before the launch ships. |
]

---

# Going Deeper

.small[
| Paper | What it adds |
|---|---|
| Goodman-Bacon (2021), *J. Econometrics* | The decomposition on these slides: exact weights, VWATT vs $\Delta$ATT estimand reading. |
| de Chaisemartin and D'Haultfoeuille (2020), *AER* | Implicit ATT(g,t) weights for general two-way FE designs; can be negative; robust switcher estimator. |
| Borusyak, Jaravel and Spiess (2024), *ReStud* | Imputation view: efficiency under parallel trends, the "forbidden extrapolation" framing. |
| Baker, Larcker and Wang (2022), *JFE* | How much this matters in published finance DiDs: a lot. |
| Roth, Sant'Anna, Bilinski and Poe (2023), *J. Econometrics* | Survey tying the whole modern DiD literature together. |
]

**Next module:** the estimators that fix this: Callaway-Sant'Anna,
Sun-Abraham, BJS imputation, dCDH.

**Drill:** `exercise.R` hand-codes all nine 2x2s and their weights, and
replicates `bacon()` to 1e-14.

---

class: center, middle, inverse

# Backup Slides

---

name: staggered-dgp

# Backup: The DGP

.small[
```{r dgp-show, eval=FALSE}
make_panel <- function(scenario = c("constant", "dynamic", "heterogeneous"),
                       n_cities = 30, n_t = 60, seed = 42) {
  scenario <- match.arg(scenario)
  set.seed(seed)
  cohorts <- tibble(
    city = 1:n_cities,
    g = c(rep(15, 8), rep(25, 8), rep(35, 8), rep(Inf, 6)) # adoption week
  )
  expand_grid(city = 1:n_cities, t = 1:n_t) |>
    left_join(cohorts, by = "city") |>
    mutate(
      treated = t >= g,
      eff = case_when(
        !treated ~ 0,
        scenario == "constant"      ~ 1.0,
        scenario == "dynamic"       ~ 0.4 * (1 + 0.10 * (t - g)),
        scenario == "heterogeneous" ~ 0.4 * (1 + 0.10 * (t - g)) *
                                      (1 + 0.8 * (g == 15) - 0.8 * (g == 35))
      ),
      y = 5 + 0.05 * t + city * 0.1 + eff + rnorm(n(), 0, 0.3) # common trend
    )
}
```
]

.small[
- City FE ( $0.1 \times$ city index) and a common weekly trend $0.05t$: parallel
  trends holds **by construction**. Every bias on the main slides is purely the
  estimator, never a PT violation.
- Noise $\sigma = 0.3$ is small relative to effects, so decomposition tables are
  stable and readable.
]

<a href="#staggered-dgp-main" class="nav-btn-br">← back</a>

---

name: fwl-detail

# Backup: The FWL View of TWFE

.small[
Frisch-Waugh-Lovell: the TWFE coefficient equals OLS of $y$ on double-demeaned
treatment,

$$\tilde D_{it} = D_{it} - \bar D_i - \bar D_t + \bar{\bar D}$$

$$\hat\beta^{DD} = \frac{\sum_{it} \tilde D_{it}\, y_{it}}{\sum_{it} \tilde D_{it}^2}$$

Three consequences:

1. $y$ enters linearly with weights $\tilde D_{it} / \sum \tilde D^2$: TWFE is a
   fixed linear combination of cell means, which is why an exact decomposition
   into 2x2s exists at all.
2. A unit treated for the entire panel has $D_{it} = \bar D_i = 1$, so
   $\tilde D_{it} \approx 0$: always-treated units contribute *nothing* to
   identification, they only sharpen the FE estimates.
3. $\sum \tilde D_{it}^2$ is maximized for units treated half the panel: this is
   exactly the $\bar D_k (1 - \bar D_k)$ term in the Bacon weights.

The decomposition then follows from partitioning $\sum \tilde D \tilde y$ by
timing-group pairs and completing each piece into a 2x2 DiD (next backup).
]

<a href="#twfe-refresher-main" class="nav-btn-br">← back</a>

---

name: bacon-derivation

# Backup: Deriving the Decomposition (1 of 2)

.small[
Setup: timing groups $a$ (earlier, adoption $g_a$), $b$ (later), $U$ (never).
Within any **pair** of groups, restricted to the right **window**, the panel is
a plain 2x2: one group changes status, the other does not.

**Step 1: partition the identifying variance.** Write the total demeaned
variance $\widehat{Var}(\tilde D)$ as a sum over group pairs and windows.
Every $(i,t)$ cell belongs to some pair-window combination, and within each,
demeaned treatment has the 2x2 structure.

**Step 2: each piece is a 2x2 DiD.** For a pair $(k, U)$, OLS on the subsample
is textbook DiD:

$$\hat\beta_{kU} = \left[\bar y_k^{post(k)} - \bar y_k^{pre(k)}\right] -
\left[\bar y_U^{post(k)} - \bar y_U^{pre(k)}\right]$$

For timing pairs $(k, l)$ with $g_k < g_l$, the window splits in two:

- $t < g_l$: only $k$ switches. Control = $l$ (not yet treated). This is
  $\hat\beta_{kl}^{k}$.
- $t \geq g_k$: only $l$ switches. Control = $k$ (treated throughout). This is
  $\hat\beta_{kl}^{l}$, the forbidden one.
]

<a href="#bacon-thm-main" class="nav-btn-br">← back</a>

---

count: false

# Backup: Deriving the Decomposition (2 of 2)

.small[
**Step 3: reassemble.** OLS on the full sample equals the variance-weighted
average of the subsample OLS coefficients:

$$\hat\beta^{DD} = \sum_{\text{pieces } p} \underbrace{\frac{\widehat{Var}_p(\tilde D)\, n_p}{\sum_q \widehat{Var}_q(\tilde D)\, n_q}}_{s_p} \hat\beta_p$$

Computing $\widehat{Var}_p(\tilde D)$ for each piece type gives the formulas on
the weights slide:

- Within a $(k,U)$ piece, treatment share is $n_{kU}$ across units and
  $\bar D_k$ across time: variance $n_{kU}(1-n_{kU}) \bar D_k(1-\bar D_k)$.
- The timing pieces get the analogous terms with the window lengths
  $(1 - \bar D_l)$ or $\bar D_k$ scaling how much panel they occupy: the
  squared leading factors.

**Why weights are positive:** they are variances times sample shares. All the
sign trouble lives inside the $\hat\beta_{kl}^{l}$ *estimates*, not the weights.

Full details: Goodman-Bacon (2021), Theorem 1 and Appendix A. The exercise
implements exactly these formulas and matches `bacon()` to 1e-16 on weights.
]

<a href="#bacon-thm-main" class="nav-btn-br">← back</a>

---

name: weights-detail

# Backup: Reading the Weight Formulas

.small[
$$s_{kU} \propto (n_k + n_U)^2\, n_{kU}(1-n_{kU})\, \bar D_k(1-\bar D_k)$$

Three factors, three design questions:

| Factor | Question it answers |
|---|---|
| $(n_k + n_U)^2$ | how big is this subsample relative to the panel? |
| $n_{kU}(1 - n_{kU})$ | how balanced are treatment and control *groups*? |
| $\bar D_k (1 - \bar D_k)$ | how balanced are pre and post *periods*? |

For timing pairs, the same three factors appear, but the window is shorter:
the leading factor becomes $\left[(n_k+n_l)(1-\bar D_l)\right]^2$ for the
early-vs-late piece (window is $l$'s pre-period, length share $1-\bar D_l$)
and $\left[(n_k+n_l)\bar D_k\right]^2$ for the forbidden piece (window is
$k$'s post-period, length share $\bar D_k$).

**Worked example from the deck's panel:** cohort g=25 spends
$\bar D = 36/60 = 0.6$ of the panel treated, closest of the three cohorts to
$0.5$, and the never-treated group is large, so $s_{25,U} = 0.18$ is the
biggest single vs-never weight even though all cohorts have 8 cities.

Consequence worth saying in an interview: **two rollouts with identical
effects but different calendars give different TWFE estimands.**
]

<a href="#bacon-weights-main" class="nav-btn-br">← back</a>

---

name: forbidden-detail

# Backup: The Forbidden 2x2, Algebraically

.small[
Take the pair early $k$ (adopts $g_k$), late $l$ (adopts $g_l > g_k$), window
$t \geq g_k$ where $k$ is treated throughout. Split the window at $g_l$:
$W_1 = \lbrack g_k, g_l)$ and $W_2 = \lbrack g_l, T\rbrack$. With parallel
trends in untreated outcomes and writing $\text{ATT}_k(W)$ for cohort $k$'s
average effect over window $W$:

$$\hat\beta_{kl}^{l} \;\xrightarrow{p}\;
\underbrace{\text{ATT}_l(W_2)}_{\text{what you want}} \;-\;
\underbrace{\left[\text{ATT}_k(W_2) - \text{ATT}_k(W_1)\right]}_{\text{control group's effect drift}}$$

- **Constant effects:** $\text{ATT}_k(W_2) = \text{ATT}_k(W_1)$, the drift term
  dies, and the forbidden 2x2 is a fine estimate of $\text{ATT}_l$.
- **Dynamic effects:** $k$'s effect keeps growing into $W_2$, the drift term is
  positive and gets subtracted: attenuation, and with strong dynamics, sign
  reversal.

In the deck's dynamic panel: the (35 vs 15) forbidden estimate is $-0.005$
while the true $\text{ATT}_{35}(W_2) \approx 0.9$. The early cohort's drift
over weeks 35-60 was that large.

This is also why the *constant* scenario is immune despite identical weights:
weights did not change, the drift terms all vanished.
]

<a href="#forbidden-main" class="nav-btn-br">← back</a>

---

name: estimand-detail

# Backup: The Estimand Decomposition

.small[
Take plims under parallel trends for untreated potential outcomes. Goodman-Bacon
shows

$$\text{plim}\; \hat\beta^{DD} = \text{VWATT} + \text{VWCT} - \Delta\text{ATT}$$

| Term | Meaning | Dies when |
|---|---|---|
| VWATT | variance-weighted ATT: each 2x2's ATT weighted by the Bacon $s$ | never (this is the "good" part) |
| VWCT | variance-weighted common trend: PT violations across pairs | parallel trends holds |
| $\Delta$ATT | weighted sum of *within-cohort effect changes* subtracted by forbidden 2x2s | effects constant in event time |

Three readings:

1. Even with PT and constant effects, TWFE targets **VWATT**, not the ATT: the
   variance weighting is a different aggregation than "average over treated
   cells". With constant effects they coincide.
2. Dynamics enter *only* through $\Delta$ATT, and always with a minus sign:
   dynamic effects bias TWFE **toward zero and beyond**, never away from zero.
3. PT violations and heterogeneity bias are *separate terms*: fixing one does
   not fix the other. Module 2 fixes $\Delta$ATT; Module 3 stress-tests VWCT.
]

<a href="#bacon-validate-main" class="nav-btn-br">← back</a>
