---
title: "Module 7: Policy Learning"
subtitle: "From HTE Estimates to Deployment Rules"
output:
  xaringan::moon_reader:
    css: [default, default-fonts]
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
---

```{css, echo=FALSE}
.remark-code, .remark-inline-code { font-size: 80%; }
.remark-slide-content { padding: 1em 2em; }
.small { font-size: 80%; }
.tiny { font-size: 65%; }
.highlight-box {
  background: #fff3e0;
  border-left: 4px solid #e65100;
  padding: 0.5em 1em;
  margin: 0.5em 0;
}
.blue-box {
  background: #e3f2fd;
  border-left: 4px solid #1565c0;
  padding: 0.5em 1em;
  margin: 0.5em 0;
}
.nav-btn {
  position: absolute;
  bottom: 12px;
  left: 40px;
  font-size: 11px;
  background: #e8eaf6;
  padding: 2px 8px;
  border-radius: 3px;
  z-index: 100;
  text-decoration: none;
  color: #1a237e;
}
.nav-btn:hover { background: #c5cae9; }
.nav-btn-br {
  position: absolute;
  bottom: 12px;
  right: 70px;
  font-size: 11px;
  background: #e8eaf6;
  padding: 2px 8px;
  border-radius: 3px;
  z-index: 100;
  text-decoration: none;
  color: #1a237e;
}
.nav-btn-br:hover { background: #c5cae9; }
.inline-btn {
  font-size: 11px;
  background: #e8eaf6;
  padding: 2px 8px;
  border-radius: 3px;
  text-decoration: none;
  color: #1a237e;
  margin-right: 6px;
  vertical-align: middle;
}
.inline-btn:hover { background: #c5cae9; }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, fig.align = "center",
  fig.width = 10, fig.height = 4.5, dpi = 150,
  message = FALSE, warning = FALSE
)
library(tidyverse)
library(grf)
library(policytree)
theme_set(theme_minimal(base_size = 14))
set.seed(42)
```

```{r make-hte, include=FALSE}
make_hte_data <- function(n = 6000, seed = 42) {
  set.seed(seed)
  tibble(
    density  = runif(n),                        # city density percentile
    tenure   = pmin(rexp(n, 1 / 18), 90),       # months on platform
    peak_shr = rbeta(n, 2, 3),                  # share of hours in peak
    rating   = pmin(pmax(rnorm(n, 4.7, 0.2), 3.5), 5),
    w        = rbinom(n, 1, 0.5),
    tau      = 1.5 * density - 0.02 * tenure + 2.0 * density * peak_shr,
    y        = 10 + 5 * density + 0.05 * tenure + 3 * peak_shr +
               tau * w + rnorm(n, 0, 2)
  )
}
```

```{r compute-all, include=FALSE, cache=FALSE}
# fit everything once; reuse across slides
dat  <- make_hte_data(6000)
cost <- 0.6
covs <- c("density", "tenure", "peak_shr", "rating")
X <- as.matrix(dat[, covs]); Y <- dat$y; W <- dat$w; e <- 0.5

cf      <- causal_forest(X, Y, W, W.hat = e, num.trees = 2000, seed = 42)
tau_hat <- predict(cf)$predictions
Gamma   <- double_robust_scores(cf)          # n x 2: control, treated

Gamma_net <- Gamma                            # cost-aware: net the treated col
Gamma_net[, "treated"] <- Gamma_net[, "treated"] - cost
tree      <- policy_tree(X, Gamma_net, depth = 2)
tree_free <- policy_tree(X, Gamma,     depth = 2)   # cost-free, for contrast
act       <- predict(tree, X)                 # 1 = control, 2 = treat

opt_share    <- mean(dat$tau  > cost)
plugin_share <- mean(tau_hat  > cost)
tree_share   <- mean(act == 2)

# honest 50/50 split for off-policy evaluation (fast)
set.seed(1)
tr <- sample(c(TRUE, FALSE), nrow(X), replace = TRUE, prob = c(.5, .5)); te <- !tr
cf_tr <- causal_forest(X[tr, ], Y[tr], W[tr], W.hat = e, num.trees = 2000, seed = 1)
yf_tr <- regression_forest(X[tr, ], Y[tr], num.trees = 2000, seed = 1)
Gtr   <- double_robust_scores(cf_tr); Gtr[, "treated"] <- Gtr[, "treated"] - cost
tree_tr <- policy_tree(X[tr, ], Gtr, depth = 2)

tau_te <- predict(cf_tr, X[te, ])$predictions
m_te   <- predict(yf_tr, X[te, ])$predictions
mu1 <- m_te + (1 - e) * tau_te; mu0 <- m_te - e * tau_te
Gte <- cbind(control = mu0 + (1 - W[te]) / (1 - e) * (Y[te] - mu0),
             treated = mu1 +      W[te]  /      e  * (Y[te] - mu1))

dr_val <- function(a) {                        # a in {1,2}; net of cost
  ch <- ifelse(a == 2, Gte[, "treated"] - cost, Gte[, "control"])
  c(value = mean(ch), se = sd(ch) / sqrt(length(ch)))
}
policies <- list(
  `treat none`    = rep(1L, sum(te)),
  `treat all`     = rep(2L, sum(te)),
  `oracle`        = ifelse(dat$tau[te] > cost, 2L, 1L),
  `forest plug-in`= ifelse(tau_te      > cost, 2L, 1L),
  `depth-2 tree`  = predict(tree_tr, X[te, ]))
base_val <- dr_val(policies[["treat none"]])["value"]
league <- imap_dfr(policies, ~ tibble(policy = .y,
  value = dr_val(.x)["value"], se = dr_val(.x)["se"],
  gain = dr_val(.x)["value"] - base_val, share = mean(.x == 2)))

# naive same-data value of the learned tree (the optimism)
a_naive <- predict(tree_tr, X[tr, ])
Gtr_full <- double_robust_scores(cf_tr)
naive_gain <- mean(ifelse(a_naive == 2, Gtr_full[, "treated"] - cost,
                          Gtr_full[, "control"])) - mean(Gtr_full[, "control"])
honest_tree_gain <- league$gain[league$policy == "depth-2 tree"]
```

# Course Map

<table>
<tr><th>#</th><th>Module</th><th>Status</th></tr>
<tr><td>1</td><td><a href="../module-01/slides.html">TWFE Diagnosed: Goodman-Bacon and the Zoo of 2x2s</a></td><td>done</td></tr>
<tr><td>2</td><td><a href="../module-02/slides.html">Heterogeneity-Robust DiD: CS, SA, BJS, dCDH</a></td><td>done</td></tr>
<tr><td>3</td><td><a href="../module-03/slides.html">Honest DiD: Sensitivity Bounds for Parallel Trends</a></td><td>done</td></tr>
<tr><td>4</td><td><a href="../module-04/slides.html">Synthetic Control: Estimator, Inference, Variants</a></td><td>done</td></tr>
<tr><td>5</td><td><a href="../module-05/slides.html">Synthetic DiD and the Bridge from SC to DiD</a></td><td>done</td></tr>
<tr><td>6</td><td><a href="../module-06/slides.html">Causal Forest: Honest Splitting and Asymptotics</a></td><td>done</td></tr>
<tr><td><b>7</b></td><td><b>Policy Learning: From HTE to Deployment Rules</b> <i>(you are here)</i></td><td>current</td></tr>
<tr><td>8</td><td><a href="../module-08/slides.html">Matrix Completion and the Modern Panel Toolbox</a></td><td>done</td></tr>
</table>

---

# This Module

Module 6 gave us $\hat\tau(x)$: a calibrated map from driver features to
conditional effects. That answers *to whom* the push works. It does **not**
tell us what to deploy.

--

**You will be able to:**

1. State the policy learning problem: choose $\pi: X \to \{0,1\}$ in a class
   $\Pi$ to maximize the value $V(\pi)$, and say why welfare maximization is a
   different objective from CATE estimation.
2. Build doubly-robust scores and run empirical welfare maximization; state
   the Athey-Wager (2021) regret bound and why restricting $\Pi$ helps.
3. Fit `policy_tree`, read the rule, and make it cost-aware.
4. Evaluate a policy off-policy with cross-fitting, and avoid the naive trap
   of scoring a policy on its own training data.

--

**Running application:** the zone-notification push to 6000 drivers from
Module 6, now with a per-push cost $c = 0.6$. Outcome: weekly completed
trips (read: engagement or revenue per customer in a retail setting). When
to push, and to whom?

---

# The Same Problem at an Online Retailer

- The business decision: given a fixed promotional budget, which customers receive the signup discount? The goal is to maximize incremental membership revenue, not merely the average discount effect.
- `policy_tree` turns $\hat\tau(x)$ into a shallow decision tree splitting on tenure and pre-period spend, producing a deployable targeting rule without requiring a per-customer score lookup.
- Naive targeting on in-sample uplift is overoptimistic: the highest-predicted decile in the training data exploits noise and is rarely the highest-uplift decile at deployment.
- The doubly-robust AIPW score gives an honest off-policy value estimate: it corrects the plug-in rule's in-sample optimism and provides a reliable lift projection for the targeting policy.
- A fixed promo budget acts as a cost constraint; the policy's welfare is evaluated on a held-out test set, not on the data used to learn the rule.

---

name: problem-main

# The Policy Learning Problem

A policy is a map $\pi : X \to \{0, 1\}$: a treat / do-not-treat decision for
every covariate profile. Its **value** is the expected outcome under
deployment:

$$V(\pi) = E\big[ Y(\pi(X)) \big] = E\big[ Y(0) + \pi(X)\,\tau(X) \big].$$

--

We want the best rule in a class $\Pi$:

$$\pi^\star = \arg\max_{\pi \in \Pi}\, V(\pi).$$

- Unconstrained optimum: the pointwise rule $\pi(x) = \mathbb{1}\{\tau(x) > c\}$
  for a per-unit cost $c$.
- Restricted $\Pi$ (shallow trees, monotone or linear-index rules): a little
  value traded for interpretability, fairness, implementability.

--

.blue-box[
ATE answered "ship?". HTE answered "to whom is it strong?". Policy learning
answers "**what rule do we deploy?**"
]

---

name: welfare-vs-estimation-main

# Welfare Maximization $\neq$ Welfare Estimation

The single most important idea in the module. Estimating $\tau(x)$ well and
choosing a good policy are **different objectives**.

.pull-left[
**CATE estimation** (Module 6)

Judged on error in *levels*:
$$E\big[(\hat\tau(x) - \tau(x))^2\big]$$
pointwise coverage, calibration.
]

.pull-right[
**Policy learning** (this module)

Judged only on getting the *sign* of
$\tau(x) - c$ right, weighted by
$|\tau(x) - c|$. Needs **ranking**, not
unbiased levels.
]

--

.highlight-box[
A biased $\hat\tau$ that preserves the ordering around $c$ gives the optimal
policy. An unbiased but noisy $\hat\tau$ can give a worse one. All the action
is near the boundary $\tau(x) = c$, exactly where levels are hardest and least
consequential for MSE.
]

--

So optimize welfare **directly** on unbiased value estimates, not a plug-in of
a CATE fit. <a href="#regret-detail" class="inline-btn">regret theory</a>

---

name: dr-scores-main

# Doubly-Robust Scores

To optimize welfare from finite data we need an unbiased, low-variance
estimate of $V(\pi)$. Build a per-arm AIPW score, an unbiased estimate of the
potential outcome under each arm:

$$\Gamma_i(w) = \hat\mu_w(X_i) +
\frac{\mathbb{1}\{W_i = w\}}{\hat e_w(X_i)}\big(Y_i - \hat\mu_w(X_i)\big),
\qquad w \in \{0, 1\}.$$

--

- $\hat\mu_w(x) = \hat E[Y \mid X = x, W = w]$: outcome model.
- $\hat e_w(x) = \hat P(W = w \mid X = x)$: propensity. Here **known**:
  $\hat e_1 = 0.5$.
- **Doubly robust:** unbiased if *either* $\hat\mu_w$ or $\hat e_w$ is right.

--

`grf` builds them from the causal forest, using its marginal fit $\hat m$ and
effect $\hat\tau$: $\hat\mu_1 = \hat m + (1-\hat e)\hat\tau$,
$\hat\mu_0 = \hat m - \hat e\hat\tau$.

```{r dr-show, eval=FALSE}
Gamma <- policytree::double_robust_scores(cf)  # n x 2: control, treated
```

<a href="#aipw-derivation" class="nav-btn">the algebra</a>

---

name: plugin-inferior-main

# Why the Plug-In Rule Is Inferior

The tempting shortcut: $\hat\pi(x) = \mathbb{1}\{\hat\tau(x) > c\}$ straight off
the forest. It is dominated by learning on DR scores.

--

1. **Regularization bias.** $\hat\tau$ is a *regularized*, biased estimate. The
   plug-in inherits that bias at the boundary, precisely where decisions flip.
   The DR score's residual term corrects first-order outcome-model error.

--

2. **Wrong loss.** The forest split criterion targets CATE heterogeneity
   (MSE-like), not welfare. Learning the policy on $\Gamma$ optimizes value
   directly and is Neyman-orthogonal: welfare is insensitive to first-order
   nuisance error.

--

.blue-box[
Use the forest to build $\hat\tau$ and $\hat\mu$; use the **scores** to choose
the rule. Not $\hat\tau > c$.
]

---

name: ewm-main

# Empirical Welfare Maximization

Athey and Wager (2021): the learned policy maximizes summed DR reward over the
class $\Pi$:

$$\hat\pi = \arg\max_{\pi \in \Pi}\, \frac{1}{n}\sum_{i=1}^n
\Big[\Gamma_i(1)\,\pi(X_i) + \Gamma_i(0)\,(1 - \pi(X_i))\Big].$$

--

Performance is **regret** against the best rule in the class:

$$R(\hat\pi) = V(\pi^\star_\Pi) - V(\hat\pi)
= O_p\!\left(\sqrt{\tfrac{\text{VC}(\Pi)}{n}}\right).$$

--

.highlight-box[
**Restricting $\Pi$ is a feature.** Regret is measured against the best rule
*in the class*, and the rate improves as the VC dimension shrinks. A depth-2
tree converges fast **and** is interpretable, auditable, deployable. You pick
$\Pi$ for constraints and get a tighter bound for free.
]

<a href="#regret-detail" class="nav-btn">utilitarian vs pointwise</a>

---

name: fit-tree-main

# `policy_tree` in Practice

`policy_tree` does an **exact** search over all depth $L$ axis-aligned trees,
maximizing summed DR reward. Not greedy CART: globally optimal on the welfare
objective.

```{r fit-live, eval=FALSE}
cf    <- causal_forest(X, Y, W, W.hat = 0.5, num.trees = 2000)
Gamma <- double_robust_scores(cf)          # n x 2: control, treated
tree  <- policy_tree(X, Gamma, depth = 2)  # exact search, 3 splits
predict(tree, X)                            # actions: 1 = control, 2 = treat
```

--

```{r tree-print, echo=FALSE}
tree_free
```

<a href="#depth-detail" class="nav-btn">depth trade-offs</a>

---

name: read-tree-main

# Reading the Rule (Cost-Aware)

Net the cost: subtract $c$ from the treated column of $\Gamma$ before fitting,
so the tree only treats where the effect clears the bar.

```{r cost-live, eval=FALSE}
Gamma_net <- Gamma
Gamma_net[, "treated"] <- Gamma_net[, "treated"] - cost   # c = 0.6
tree <- policy_tree(X, Gamma_net, depth = 2)
```

--

```{r tree-cost-print, echo=FALSE}
tree
```

--

.small[
Newer drivers (tenure below about 28 months) are treated once density clears
about 0.27; veterans only in the densest markets (density above about 0.73).
Matches the DGP: tenure lowers $\tau$, so veterans need a denser market to pay
off. The rule treats `r sprintf("%.0f%%", 100*tree_share)` of drivers.
]

---

name: cost-shifts-main

# The Cost Bends the Rule

Same DR scores, three cost levels. Higher $c$ shrinks the treat region: the
effect must clear a higher bar.

```{r cost-sweep, echo=FALSE, fig.height=3.4}
sweep <- map_dfr(c(0.0, 0.6, 1.2), function(cc) {
  G <- Gamma; G[, "treated"] <- G[, "treated"] - cc
  tt <- policy_tree(X, G, depth = 2)
  tibble(cost = cc, share = mean(predict(tt, X) == 2))
})
tibble(cost = seq(0, 1.6, by = 0.2)) |>
  mutate(opt_share = map_dbl(cost, ~ mean(dat$tau > .x))) |>
  ggplot(aes(cost, opt_share)) +
  geom_line(linewidth = 1, color = "#1565c0") +
  geom_point(data = sweep, aes(cost, share), size = 4, color = "#e65100") +
  annotate("text", x = 1.2, y = 0.72, size = 3.6, color = "#e65100",
           label = "orange = depth-2 tree share") +
  labs(x = "Per-push cost c", y = "Share treated",
       title = "True optimal share (line) and the tree's share (points) fall with cost")
```

At $c = 0.6$ the true optimal share is
`r sprintf("%.0f%%", 100*opt_share)`; the tree treats
`r sprintf("%.0f%%", 100*tree_share)`.

---

name: ope-main

# Off-Policy Evaluation

Given any candidate $\pi$, estimate its value from RCT data with the
doubly-robust value estimator:

$$\hat V(\pi) = \frac{1}{n}\sum_{i=1}^n
\Big[\Gamma_i(0) + \pi(X_i)\big(\Gamma_i(1) - \Gamma_i(0)\big)\Big],$$

net of cost by using $\Gamma_i(1) - c$ for the treated term.

--

- An average of i.i.d. terms $\psi_i$, so the standard error is the
  influence-function form $\widehat{\text{SE}} = \hat\sigma_\psi / \sqrt{n}$.
- Lets you build a **league table**: score treat-none, treat-all, the oracle,
  the plug-in, and the tree on one common yardstick, with SEs.

--

.highlight-box[
One catch. If you score a *learned* policy on the data that trained it,
$\hat V(\hat\pi)$ is optimistic: the rule adapted to that noise. Cross-fit.
]

<a href="#ope-derivation" class="nav-btn">influence function + cross-fit</a>

---

name: naive-trap-main

# The Naive Trap, Live

Same depth-2 tree, two ways to value it: on its own training scores, versus on
a held-out fold whose scores it never saw.

```{r trap-table, echo=FALSE}
tibble(
  evaluation = c("naive: same data", "honest: held-out fold"),
  `tree gain over treat-none` = c(naive_gain, honest_tree_gain)
) |>
  mutate(across(where(is.numeric), ~ round(.x, 3))) |>
  knitr::kable(format = "html")
```

--

.highlight-box[
The naive number (`r sprintf("%.2f", naive_gain)`) overstates the honest one
(`r sprintf("%.2f", honest_tree_gain)`). That gap is pure optimism: the tree
fit the noise in the scores it was trained on. A held-out evaluation, or
$K$-fold cross-fit, removes it.
]

The exercise builds the full 5-fold cross-fit. <a href="#hte-dgp" class="nav-btn">DGP code</a>

---

name: league-main

# The League Table (Honest)

Off-policy values on a held-out fold, as gains over treat-none, net of cost.

```{r league-table, echo=FALSE}
league |>
  transmute(policy,
            `gain vs none` = round(gain, 3),
            SE = round(se, 3),
            `share treated` = round(share, 3)) |>
  knitr::kable(format = "html")
```

--

.small[
- **treat-all** pays the cost everywhere: gain
  `r sprintf("%.2f", league$gain[league$policy=="treat all"])`, the weakest
  non-trivial policy.
- the infeasible **oracle** rule $\tau > c$ and the **plug-in** land near
  `r sprintf("%.2f", league$gain[league$policy=="oracle"])`.
- the **depth-2 tree** matches them at
  `r sprintf("%.2f", honest_tree_gain)`, within one SE of the oracle, using
  just three interpretable splits.
]

--

.blue-box[
The interpretable rule loses almost nothing to the infeasible oracle here, and
beats treat-all by about
`r sprintf("%.2f", league$gain[league$policy=="depth-2 tree"] - league$gain[league$policy=="treat all"])`
trips per driver net of cost. That is the case for policy learning over
blanket rollout.
]

---

name: failure-modes-main

# When Policy Learning Breaks

.small[
| Failure | Symptom | What to do |
|---|---|---|
| Bad nuisances | DR score variance explodes; wild $\hat V$ | check overlap, trim propensities; DR fixes first-order error, not gross misspecification |
| Naive evaluation | learned policy "beats" the oracle on its own data | cross-fit or honest split, always |
| Over-rich $\Pi$ | deep tree, slow regret, no interpretability | depth 2 default; exact search cost grows fast with depth |
| Boundary-heavy DGP | small honest gain, wide SEs | report the SE; no rule separates much from treat-all |
]

--

.highlight-box[
Interview framing: policy learning is not "predict $\tau$ then threshold". It
is "optimize deployment welfare directly on doubly-robust scores, over a class
you can actually ship, and evaluate it honestly with cross-fitting".
]

--

.tiny[
**Two boundary cases.** (a) Offline vs online: a stationary environment and
a one-shot targeting rule call for learning offline from logged data, this
module; cheap iteration and fast feedback call for a bandit that explores
for you online. (b) A budget constraint ("treat at most $B$ units") turns
$\tau(x) > c$ into "treat the top $B$ by score", connecting straight back to
Module 6's TOC/RATE machinery.
]

---

name: m7-interview-questions

# Interview Questions

.small[
| Question | Core of a strong answer |
|---|---|
| "We have a good CATE model. Why not just push when $\hat\tau > c$?" | Plug-in inherits regularization bias at the boundary and optimizes MSE, not welfare. Learn the rule on doubly-robust scores: Neyman-orthogonal, corrects first-order nuisance error. |
| "Why restrict to a depth-2 tree instead of the best possible rule?" | Regret is against the best rule in the class and improves as VC dimension shrinks; a shallow tree converges fast and is interpretable, auditable, deployable. Restriction is a feature. |
| "How do you know the deployed policy is actually good?" | Off-policy DR value estimate with SEs, cross-fitted so the policy is never scored on its training data. Report gain over treat-all and treat-none. |
| "Your policy beat the oracle in the eval. Believe it?" | No. That is the optimism of same-data evaluation. Cross-fit; the honest value drops below the oracle. |
| "The gain over treat-all is tiny. So what?" | Most units may sit near $\tau = c$; then no rule separates much value. Report the SE, and weigh interpretability and cost savings, not just the point estimate. |
| "Finance says every push costs money. Can we just stop sending the useless ones?" | That is exactly a cost-aware policy: net the cost out of the doubly-robust scores, learn the rule on the net scores, and report the cross-fitted value against treat-all and treat-none before shipping it. |
]

---

# Going Deeper

.small[
| Paper | What it adds |
|---|---|
| Athey and Wager (2021), *Econometrica* | The framework on these slides: EWM with doubly-robust scores, the $\sqrt{\text{VC}/n}$ regret bound, observational-data guarantees. |
| Kitagawa and Tetenov (2018), *Econometrica* | Empirical welfare maximization from experimental data; the original regret-based treatment-choice result. |
| Zhou, Athey and Wager (2023), *Operations Research* | Multi-action policy learning; the `policytree` exact-search algorithm and its optimization. |
| Wager and Athey (2018), *JASA* | The causal forest supplying $\hat\tau$ and $\hat\mu$ for the scores. |
| Hirano and Porter (2009), *Econometrica* | Asymptotic optimality of treatment rules; the decision-theoretic backdrop. |
]

**This closes the arc:** estimand (M1-3), design (M4-5), heterogeneity (M6),
deployment (M7). Module 8 returns to panels with matrix completion.

**Drill:** `exercise.R` fits the tree, hand-builds the 5-fold cross-fitted DR
value estimator, and shows the optimism gap plus the honest league table.

---

class: center, middle, inverse

# Backup Slides

---

name: aipw-derivation

# Backup: The AIPW Score and Double Robustness

.small[
The score for arm $w$ is an outcome-model prediction plus an
inverse-propensity-weighted residual:

$$\Gamma_i(w) = \hat\mu_w(X_i) +
\frac{\mathbb{1}\{W_i = w\}}{\hat e_w(X_i)}\big(Y_i - \hat\mu_w(X_i)\big).$$

**Unbiased if the propensity is right.** Take $E[\cdot \mid X_i]$ with the true
$e_w$: the residual term has mean $e_w \cdot E[Y - \mu_w \mid X, W = w] / e_w$,
so even a wrong $\hat\mu_w$ is corrected back to $E[Y(w) \mid X_i]$.

**Unbiased if the outcome model is right.** If $\hat\mu_w = \mu_w$, the residual
has conditional mean zero regardless of $\hat e_w$, leaving $\mu_w(X_i)$.

So the score is consistent if *either* nuisance is correct: double robustness.
More: the value estimate is **Neyman-orthogonal**, its first-order derivative
in each nuisance is zero, so small nuisance errors do not bias welfare. That is
why we learn the policy on $\Gamma$ rather than on $\hat\tau$ directly.

`grf` construction from the causal forest: with marginal fit $\hat m(x)$,
effect $\hat\tau(x)$, propensity $\hat e$,
$\hat\mu_1 = \hat m + (1-\hat e)\hat\tau$ and
$\hat\mu_0 = \hat m - \hat e\hat\tau$.
]

<a href="#dr-scores-main" class="nav-btn-br">← back</a>

---

name: regret-detail

# Backup: Regret, VC, and the Pointwise Impossibility

.small[
$$R(\hat\pi) = V(\pi^\star_\Pi) - V(\hat\pi)
= O_p\!\left(\sqrt{\tfrac{\text{VC}(\Pi)}{n}}\right).$$

The rate comes from a uniform bound on the empirical welfare process over
$\Pi$: with DR scores the objective is a sum of mean-zero-error terms, and its
supremum over a class of VC dimension $d$ deviates from the truth by
$O_p(\sqrt{d/n})$. Depth $L$ trees on $p$ covariates have VC dimension growing
with the number of leaves, so shallow trees give the tightest bounds.

**Utilitarian, not pointwise.** No estimator recovers the pointwise-optimal
rule from finite data: near $\tau(x) = c$ you cannot resolve the sign of
$\tau(x) - c$. But the misclassified units there have $|\tau(x) - c|$ small, so
they cost little welfare. Regret weights each mistake by exactly that value
gap, so it downweights the unavoidable errors and rewards getting the
high-stakes regions right.

**Reading the bound for practice.** Doubling $n$ shrinks regret by $\sqrt 2$;
halving VC dimension (depth 3 to depth 2) shrinks it by the same factor and
buys interpretability. This is why the deployment-constrained class is not a
compromise: it is often the statistically preferable choice too.
]

<a href="#ewm-main" class="nav-btn-br">← back</a>

---

name: ope-derivation

# Backup: The DR Value Estimator and Cross-Fitting

.small[
For a fixed policy $\pi$, the value estimator averages per-unit influence
terms:

$$\hat V(\pi) = \frac{1}{n}\sum_i \psi_i, \qquad
\psi_i = \Gamma_i(0) + \pi(X_i)\big(\Gamma_i(1) - \Gamma_i(0)\big).$$

Since $E[\Gamma_i(w) \mid X_i] = E[Y(w) \mid X_i]$, we have
$E[\psi_i] = V(\pi)$, and because the $\psi_i$ are i.i.d.,
$\widehat{\text{SE}} = \hat\sigma_\psi / \sqrt{n}$. Comparisons of two policies
use the paired differences $\psi_i^{(a)} - \psi_i^{(b)}$, whose SE is usually
much smaller than the two marginal SEs (the nuisance noise cancels).

**Why cross-fit for a learned $\pi$.** If $\hat\pi$ is fit on the same scores
used to evaluate it, $\psi_i$ and $\hat\pi$ are dependent, and $E[\psi_i]$ no
longer equals $V(\hat\pi)$: the estimate is optimistic. $K$-fold cross-fit
restores independence:

1. fold $k$ held out; fit nuisances and learn $\hat\pi_{-k}$ on the other folds;
2. score fold $k$ with DR scores built from the *training-fold* nuisances;
3. evaluate $\hat\pi_{-k}$ on fold $k$; pool the held-out $\psi_i$.

Each held-out unit is scored by a policy and nuisances that never saw it, so
the pooled average is honest. The exercise implements exactly this loop.
]

<a href="#ope-main" class="nav-btn-br">← back</a>

---

name: depth-detail

# Backup: Depth Trade-Offs and Exact Search

.small[
`policy_tree` solves an **exact** optimization: over all axis-aligned trees of
depth $L$, find the one maximizing $\sum_i \Gamma_i(\pi(X_i))$. Unlike greedy
CART it does not split one node at a time; it globally optimizes, so the tree
is the true welfare maximizer in the depth $L$ class.

The price is combinatorial cost. Exact search scales with the number of
candidate split points across covariates, and grows sharply with depth:

| Depth | Leaves | Splits | Typical use |
|---|---|---|---|
| 1 | 2 | 1 | one-variable rule, sanity check |
| 2 | 4 | 3 | practical default, interpretable |
| 3 | 8 | 7 | slow on wide data, rarely worth it |

Practical guidance:

- Start at depth 2. On $n = 6000$ with four covariates it fits in seconds.
- Coarsen continuous covariates (`policytree` supports pre-binning) to speed
  the search without meaningful welfare loss.
- If depth-3 barely beats depth-2 on the cross-fitted value, keep depth 2:
  cheaper, faster, and lower regret by the VC argument.
]

<a href="#fit-tree-main" class="nav-btn-br">← back</a>

---

name: hte-dgp

# Backup: The HTE DGP

.small[
```{r dgp-show, eval=FALSE}
make_hte_data <- function(n = 6000, seed = 42) {
  set.seed(seed)
  tibble(
    density  = runif(n),                        # city density percentile
    tenure   = pmin(rexp(n, 1 / 18), 90),       # months on platform
    peak_shr = rbeta(n, 2, 3),                  # share of hours in peak
    rating   = pmin(pmax(rnorm(n, 4.7, 0.2), 3.5), 5),  # nuisance covariate
    w        = rbinom(n, 1, 0.5),               # randomized push, e = 0.5
    tau      = 1.5 * density - 0.02 * tenure + 2.0 * density * peak_shr,
    y        = 10 + 5 * density + 0.05 * tenure + 3 * peak_shr +
               tau * w + rnorm(n, 0, 2)         # weekly completed trips
  )
}
```

- Shared with Module 6, so $\hat\tau$ and the league table are comparable
  across the two decks.
- True effect $\tau(x)$ rises with density and peak share, falls with tenure;
  `rating` has no effect (a decoy the tree should ignore).
- Cost $c = 0.6$ per push. True optimal share treated, $\tau > 0.6$, is about
  `r sprintf("%.0f%%", 100*opt_share)`.
]

<a href="#naive-trap-main" class="nav-btn-br">← back</a>
